### 2024-02-12 maksakov-readings-conf-2024/Conference-paper.Rmd
Minor fixes in CSL translation and docx styles, compression of paper text, finish of translation of plots and tables, fix of a bug with paths table

---
title: "UDC 911.3"
subtitle: "Migration of Russian Legal Companies in 2016–23"
author: P. O. Syomin
output: 
  word_document:
    reference_docx: conference-paper-reference.docx
bibliography: refs.bib
csl: "gost-r-7-0-5-2008-numeric-alphabetical-en.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 6.4, 
  fig.height = 5,
  fig.id = FALSE,
  dpi = 300)
library(knitr)
```

```{r imports, include=FALSE, cache=TRUE}
source("Script.R")
```

<div custom-style="Date">

Perm State University, 15 Bukireva st., Perm, 614068, Russia, ntsp\@ya.ru

</div>

<div custom-style="Abstract">

Migration of Russian commercial legal companies in 2016–23 is described and analysed. Distributions by migration count and distance are estimated and visualized, the relationship between company size and migration decision is statistically tested, net migration rates by region are mapped, and top migration paths are listed. It is found that migrations are rare, mostly short-distance (10–50 km). Bigger companies tend to relocate more often. Moscow has probably lost small and inefficient businesses but gained larger and more profitable.

<br>

***Keywords:*** companies migration, small and medium-sized businesses, legal companies, legal service, service geography, data analysis, open data, Federal Tax Service.

</div>

<br>

Migration of companies (business relocation) is an important phenomenon of economy and geography. Its first studies date back to the middle of 20th century, starting from G. McLaughlin and S. Robock's book *Why industry moves South* that have described the migration of enterprises from the North of the US and linked it to the cheaper labor and smaller activity of labor unions [@McLaughlin1949]. Various aspects of companies migration have been studied, including factors (reasons) [@Wu2007], directions (places) [@Rossi2019], taxation effects [@Kubicova2016].

In Russia, there are few papers about migration of companies. Existing works focus  on relocation as an instrument of tax optimization [@Adigamova2011_en; @Shcherbakova2016_en], skipping the geographical overview and spatial aspects of the phenomenon. So, a spatial analysis of business relocation in Russia is important. This study describes and visualizes the migration of Russian commercial legal companies in 2016–23. The choice of particular area of economy, that is, legal service, is subjective.

**Data.** A geo-referenced dataset of Russian small and medium-sized businesses (SMBs) with main activity code equal to 69.10 (activity in the area of law) was generated from the open dumps of SMB registry[^1] with an open source Python CLI application *ru-smb-companies*[^2] developer by the author of this paper. The dataset covers the period from August 2016 to November 2023. According to [@Moiseeva2016_en], by August 2016 in Russia there have been 47000 commercial legal companies, and the dataset contains about `r round(companies_count_08_2016, -3)` companies with the same date. Some legal companies are missing, e. g. large ones that exceed the limit on the revenue or number of employees for SMBs, but 3/4 are present, so the dataset looks reliable and relevant.

[^1]: https://www.nalog.gov.ru/opendata/7707329152-rsmp/
[^2]: https://github.com/PavelSyomin/russian-smb-companies

**Methods.** Company migration is seen as a change in address of registration [@Shcherbakova2016_en]. It is narrow and does not account for other ways of relocation (e. g. mergers or reorganization). Also, it is formal, because the change of nominal address does not necessarily reflects the change of actual place of doing business. However, it looks like the best choice for the data available. Descriptive statistics, statistical tests, data visualization and mapping are used. Technically, the analysis was performed with R programming language (including additional packages). The source code for the analysis is stored in public repository[^3].

[^3]: https://github.com/PavelSyomin/tax-service-opendata-paper/tree/main/maksakov-readings-conf-2024

A **distribution of companies by the number of migrations** is shown on *Fig. 1*. Almost all the companies (97%) in 2016–23 have not migrated. About 2.5% of companies have changed the location exactly once, and more frequent migrations are extremely rare. The distribution is almost log10-linear.

<br>

```{r fig-1-mig-distr, fig.cap="*Fig. 1.* The distribution of companies by the number of migrations. Y axis is log10-scaled. The counts are shown above the bars, and percetrages are below the top border", fig.height=2.5}
mig_count_distr_plot
```

<br>

A **relationship between the migration decision and the size of company** is depicted on *Fig. 2*. Migration among small firms is more frequent in comparison to micro-businesses: according to Fisher's exact test, `r ifelse(categories_test$p.value < 0.001, "p < 0.001", paste("p = ", categories_test$p.value))`, and odds ratio is `r paste0("1 to ", round(1 / categories_test$estimate, 0))`.

<br>

```{r fig-2-mig-by-company-size, fig.cap="*Fig. 2*. Migration decision depending on company's size (category)", fig.height=1}
mig_by_category_plot
```

<br>

A **distribution of migration distance** is shown on *Fig. 3*. The majority of businesses moves to small distance (10–50 km), but there is a significant group of long-migrating firms (the distance is about 1000 km).

<br>

```{r fig-3-distance, fig.cap="*Fig. 3.* Distribution of migration distance", fig.height=2.5}
mig_distance_plot
```

<br>

**Net migration rate by region** is shown on map (see *Fig. 4*). Most of the regions have net migration rate close to zero. However, there are some regions with substantial growth (Moscow, St. Petersburg, Krasnodar, Rostov, Novosibirsk oblast, Perm krai) or decline (Moscow as a city) in the number of registered legal companies. The absolute counts are small (<100) compared to the quantity of registered companies.

<br>

```{r fig-4-mig-by-region-plot, fig.cap="*Fig. 4.* Net migration rate of legal companies by regions", fig.height=3}
mig_by_region_plot
```

<br>

**The most popular migration paths** (by the count of migrated companies) with summed up revenues and employee counts are listed in the **Table**. The first two paths represent the relocation from two largest cities of Russia to the surrounding regions. Net change in the number of businesses is relatively small. An increase in the number of registered businesses is not necessarily associated with the increase in income or labor force, and vise versa. Positive flow of companies with negative flow of revenue or labor probably means that arriving businesses are larger and more profitable than leaving.

<br>

```{r table}
kable(
  head(regional_paths_combined_en, 10),
  col.names = c("Path", "Count", "Revenue, M rub", "Employees"),
  caption = "Top 10 paths of inter-regional migration of Russian legal companies in 2016–23",
  digits = 0
)
```

<br>

**Conclusion.** Migration of legal companies in Russia is a rare phenomenon: only about 2.5% of all businesses in this activity area have changed their nominal location in the period of 2016–23. Small companies, in comparison to micro-businesses, are more likely to relocate: the chances are 4 times higher. Short-distance (10–50 km) migrations are more common than long-distance (~1000 km) ones. Moscow and St. Petersburg regions are at the top by net migration rate estimated by the count of relocated businesses, while Moscow have lost the highest number of companies. However, it looks like that while losing small and inefficient companies, Moscow attracts more profitable businesses.

# Bibliographical references

<div id="refs"></div>