### 2024-03-30 legal-companies-regional-specialization/Paper.Rmd
Minor improvements

---
title: "Legal Services Market in Russia: Regional Specialization"
author: "Pavel O. Syomin"
date: "PhD Student at Perm State University"
output:
  word_document:
    reference: reference.docx
bibliography: biblio.bib
csl: biblio-style.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.id = FALSE,
  fig.width = 7,
  fig.height = 4)
```

```{r imports, include=FALSE, cache = TRUE}
source("Script.R")
```

<br>

<div custom-style="Abstract">

Abstract: This paper focuses on the spatial distribution of the legal services market in Russia. The aim is to identify the most specialized regions (constituent entities) using a quantitative technique based on the relative numbers of employees. The research utilizes open data distributed by the Federal Tax Service and the Federal Statistics Service of Russia. The regional specialization is measured using two metrics: concentration and localization. High concentration indicates regions with legal services of national importance, while high localization marks regions where legal services have local importance. The study's results reveal several clusters of legal services. These clusters are likely linked to the high level of economic activity in the regions. The location of appeal and cassation courts may also partially explain the clusters.

Keywords: specialization, legal services, regions, economic geography, concentration, localization, legal market, open data.

</div>

<br>

**Introduction.** Analysis of economic concentration and regional specialization is a common research task in economic geography. It helps to understand the current state of the territory and discover opportunities for its development [@Romanov2015]. There are various methods for studying concentration and specialization, including calculating coefficients [@Bryantseva2022] and analysis of graphs [@Blanutsa2020]. In particular, the quantitative technique used in the Atlas of Economic Specialization of Russian Regions [@Atlas2021] is of high interest, because it is modern, data-driven, theoretically grounded, and reproducible. The authors of the Atlas have only analyzed a subset of sectors in Russia's economy, and some important economic areas are missing. One example of a missing sector is the legal services market.

There is a lack of geographical research on legal services in Russia except for one analytical review by the members of the Institute for the Rule of Law [@Moiseeva2016]. However, some argue that studying the spatial aspects of legal service provision is important [@Joseph1982]. This is because the development of legal services reflects both business and economic activity and also stimulates social and economic development by increasing the protection of citizens' and businesses' rights and freedoms. Therefore, this research aims to analyze the regional distribution of legal services in Russia and identify the regions that specialize in providing such services.

**Objective.** The research has a twofold aim. Firstly, it aims to identify the regions in Russia that specialize in legal services. Secondly, it aims to visualize this specialization on a tile-grid map.

**Data.** This study utilizes the open data on Russian small and medium-sized businesses (SMBs) distributed by the Federal Tax Service (FTS)[^1]. Additionally, the FTS open data on the average number of employees of organizations[^2] is used. The author extracted information about legal firms (companies and individual entrepreneurs) that existed in 2021, including their region of registration and number of employees, from the source datasets. To filter out non-legal firms, the main activity code was restricted to 69.10. The data was processed using the command-line Python tool *ru-smb-companies*[^3], developed by the author of this paper. In addition to the company's data, official statistical information from the Federal Statistics Service of Russia (Rosstat) was utilized to determine the average yearly employees count by region in 2021[^4].

[^1]: URL: https://www.nalog.gov.ru/opendata/7707329152-rsmp/ (accessed March 28, 2024).
[^2]: URL: https://www.nalog.gov.ru/opendata/7707329152-sshr2019/ (accessed March 28, 2024).
[^3]: URL: https://github.com/PavelSyomin/russian-smb-companies (accessed March 28, 2024).
[^4]: URL: https://fedstat.ru/indicator/34051 (accessed March 28, 2024).

**Methods.** The study follows the technique proposed in the Atlas [@Atlas2021] and uses two metrics to estimate regional specialization. The first metric, *concentration*, is calculated as the number of legal companies employees in the region divided by the total number of legal companies employees in all regions. The second metric, *localization*, is calculated by dividing the concentration by the ratio of the number of all employees in the region and the count of all employees in Russia. It can be said that concentration indicates the region's share of the total number of legal workers, while localization shows whether this share is greater or less than the region's share in Russia's labor force.

In certain studies, only the localization metric is used to conclude about the specialization [@Romashina2019, @Bryantseva2022]. However, the authors of the Atlas propose a two-dimensional approach that combines concentration and specialization as follows. After calculating the metrics, the regions are ranked by localization, and the top 16 regions (approximately 20%) are considered to be areas where the legal services are of local importance. Next, the regions are ranked by concentration, and the cumulative concentration is calculated. Regions with a cumulative concentration below 0.8 are identified as areas where the legal services are of national importance, that is, these regions collectively account for 80% of the country's legal workers. Some regions are present in both two groups.

To visualize the results, a tile-grid map of Russia is used. Tile-grid maps are a relatively new phenomenon in Russian cartography, primarily used in popular science and journalism. However, they have also been suggested for potential use in spatial research [@Elatskov2020]. The map is drawn using a tile-grid map template created by I. Dianov[^5] and a custom implementation in the R programming language based on Yu. Tuckachev's code[^6]. The analysis was also conducted using R, and the source code for the research and paper has been published in the author's GitHub repository[^7].

[^5]: URL: https://dianov.org/all/plitochnaya-karta-rossii/ (accessed March 28, 2024).
[^6]: URL: https://gist.github.com/tukachev/c3262242a20b3050e0459b4e3afc3a51 (accessed March 28, 2024).
[^7]: URL: https://github.com/PavelSyomin/ru-smb-companies-papers/tree/main/legal-companies-regional-specialization (accessed March 28, 2024).

**Results and discussion.** Figure 1 shows the tile-grid map of regional specialization in legal services. Several clusters are distinguishable based on the importance of legal services. These include (1) Moscow (with the surrounding Moscow region), (2) Saint Petersburg, (3) the group of southern regions (Krasnodar and Stavropol territories, Voronezh, Belgorod, and Rostov regions), (4) the Volga belt formed by Volgograd, Saratov, Samara regions, Perm territory, Tatarstan and Udmurt republics, (5) the Urals cluster with Sverdlovsk and Chelyabinsk regions as the core and Bashkir republic on the peripheries, (6) Siberian row (Tyumen, Omsk, Novosibirsk, Kemerovo regions, Altai and Krasnoyarsk territories, Irkutsk region). Nizhniy Novgorod region and Primorsky territory can be treated as single-region clusters. Some regions, such as Kaliningrad, Tver, Ivanovo, Murmansk regions, and Kamchatka territory, have legal services of local importance and are not a part of any cluster.

Further investigation is required to determine the reasons for the observed spatial patterns. However, legal services are likely concentrated in Moscow and Saint Petersburg due to their high levels of economic development and concentration. The other clusters can also be explained by their level of economic development. The same hypothesis may be true for the two single-region clusters. Most specialized regions have at least one ordinary or commercial appeals or cassation court. The concentration or localization of legal services in these regions may be partially explained by the presence of such courts, which act as “gravitational centers” for both legal services customers and providers. However, this hypothesis requires further testing.

<br>

```{r figure-1-map, fig.cap="Figure 1. Tile-grid map of Russia showing the specialization of regions in legal services"}
map
```

<br>

**Conclusion.** The concentration and localization of legal services in Russia were measured using metrics based on the number of employees. The study identified six main clusters of legal services and two single-region clusters. The result may be linked to the level of economic development of the respective regions.  The presence of an appeal or cassation court may also be connected with the region's specialization in legal services.

<br>

# References

<div id="refs"></div>