### 2024-04-15 legal-companies-geography/Paper.Rmd
Data and partially methods, #1

---
output: 
  word_document:
    reference_doc: reference.docx
bibliography: refs.bib
#csl: style.csl
link-bibliography: False
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Pavel O. Syomin[^1]**

<br>

<div custom-style="Title">

Legal Business Geography in Russia: Regional Analysis Using Tax Service Open Data

</div>

<br>

# Abstract

Text text text.

<br>

<div custom-style="Block Text">

**KEYWORDS:** w1, w2

</div>

<br>

[^1]: Perm State University, Department of Social and Economic Geography; 15 Bukireva st., Perm, 614068, Russia; e-mail: ntsp\@ya.ru

# Introduction

Geography of services is an important branch of modern economic and social geography. In Russia, it dates back to the 1960s when the pioneering research papers in this field of studies were published [@Tkachenko2016]. In foreign countries, the spatial aspects of services were treated as a part of human geography and touched in the scholarly works about retail [@Reilly1931; @Berry1967] and cities [@Brush1953] usually based on the central places theory [@Christaller1966]. A series of review research articles published in 1985–87 directly uses the term “geography of services” and cites dozens of original papers related to this topic [@Daniels1985; @Daniels1986; @Daniels1987]. Newer foreign studies in this field focus on the processes and phenomena in peripheral or smaller economies [@Hermelin2007] or on the arising and growing sectors like FinTech [@Lai2020; @Wojcik2020]. In Russia, recent research in the geography of services covers spatial properties of banking [@Shustova2015], education [@Shulgina2019], car rental [@Zyrianova2018], food service [@Preobrazhenskiy2018], service sector in general [@Lisin2020]. However, several types of services are rarely studied, e. g. business and legal ones [@Tkachenko2016].

The research on the geography of legal services in Russia is limited. To date, only an analytical review describing the legal market in Russia was published that contains a chapter on its spatial aspects [@Moiseeva2016], and yet another paper briefly addresses the role of Russia in the global legal services market without diving into the internal spatial properties of the Russian legal services itself [@Yastrebov2021]. There are also non-scientific products about Russian legal market that include its regional analysis[^2]. Despite the observed low interest to this issue from other researches, it can be argued that the geographical studies of legal services is necessary. First, legal services, as a subset of other business or producer services, play a substantial role in the development of regions and territories and in the economic growth [@Hansen1993], and sometimes act as a limiting factor, that is, the lack of professional legal advice may cause problems to entrepreneurs, especially to small ones, thus decreasing their performance [@Hitchens1997]. Next, the in-depth analysis of legal companies may give valuable insights about the business sectors they serve [@Faulconbridge2019] or increase the theoretical understanding of markets and industries in general [@Henderson2016]. Last but not least, legal services, apart from their narrow economic function, should be also treated as one of the elements of legal system in a wider sense that protect human rights and freedoms, ensure the rule of law, and act as one of the intermediaries between people, business and government that help to resolve the conflicts, increase the certainty and orderliness of social relations. This aspect of legal services is closely related to the social dimension of geography of services, and from this point of view it is not surprising that the research is also being conducted on the problem of legal services availablity [@Baxter2015].

[^2]: Анализ рынка юридических услуг в России в 2019-2023 гг, прогноз на 2024-2028 гг. Web resource: https://marketing.rbc.ru/research/41683/ (accessed 14.04.2024)

Legal services is quite a narrow and specific area of economy that does not reported separately in the official statistics, so for their analysis, other sources of information are necessary. Company-level georeferenced data is valuable for this task, because it can be easily filtered to leave the firms doing the business in the required activity area only and then aggregated by its spatial attributes to make a geographical overview. Researchers typically extract such the information from public GIS services [@Lachininsky2021] or business intelligence (BI) systems [@Moiseeva2016]. The disadvantage of this method, however, is the complexity of data preprocessing or the relatively high price of the data. Also, GIS and BI systems often do not enable access to historical data, providing real-time picture only, while the geographical studies may benefit from combining spatial and temporal properties. To overcome all these issues, an open administrative data may be used. In particular, the publicly available archives of Registry of Small and Medium-Sized Businesses (SMB registry) distributed by the Federal Tax Service (FTS) of Russia, in combination with a few other open data collections of the same governmental body, look valuable for the task in hand, and they have already been used by geographers [@Rostislav2021; @Gumenyuk2022].

The purpose of this paper is twofold. On the one hand, it aims to provide an overview of geography of legal business in Russia. The scope of the research is intentionally limited due to various organizational and technical constraints: it deals with the service providers only (the supply part of the market) and contains nothing about customers (the demand part), focuses on commercial legal companies with other market participants such as notaries, advocacy firms, non-profit legal organizations left untouched, and operates with only a portion of variables and metrics indicating the characteristics of legal business. On the other hand, the paper proposes a reproducible technique of working with FTS open data for analyzing the geography of services in Russia, particularly the services that do not reported independently from the others in the official statistics (mainly specific types of business and customer services) and thus can hardly be analyzed with the traditional economic geographers' toolkit.

<br>

# Materials and methods

**Data.** The research utilizes two primary data sources. All of them are published on the FTS website as open data.

The first dataset is called *“Registry of small and medium-sized enterprises”*[^3]. It contains full dumps of the respective registry in zipped XML files. The first dump was published in August 2016, and since then, new dumps were published every month. Each dump contains the whole registry and includes, in particular, company's names, their taxpayer IDs, activity codes according to All-Russia Classifier of Kinds of Economic Activity, address of incorporation (up to a settlement name). According to Russian laws, the Registry includes information about all commercial organizations and individual entrepreneurs that meet several criteria, the most important of which are limit on revenue (no more than 2 billion Russian rubles) and average number of employees (no more than 250). Thus, the open dumps of this registry can be treated as the most complete publicly available data source on the Russian business entities. In addition, this source has geographical reference in the form of human-readable addresses, and historical records from 2016 to now in the form of time series of monthly dumps of the registry.

[^3]: Единый реестр субъектов малого и среднего предпринимательства. Web resource: https://www.nalog.gov.ru/opendata/7707329152-rsmp/ (accessed 15.04.2024)

The second dataset is called *“Information about the average list count of employees of organisations”*[^4]. As the name suggests, it contains information about the number of company's employees. The data was initially published in 2019 and since then was updated with varied frequency, but at least yearly. As a result, the data archive includes information for the years from 2019 to 2023. For the unspecified reasons, the old versions of the dataset containing information for years from 2019 to 2021 were once deleted from the FTS website, but they had been already downloaded by the author and thus used in this research.

[^4]: Сведения о среднесписочной численности работников организации. Web resource: https://www.nalog.gov.ru/opendata/7707329152-sshr2019/ (accessed 15.04.2024)

In addition to the primary data sources, two lookup tables with the information on Russian cities and settlements are used to normalise and geocode raw addresses contained in the FTS data. The first table is called *“Settlement of Russia: population and geographical coordinates”*[^5]. It was prepared by the non-profit open data initiative called “Research data infrastructure” (RDI). The second table called *“Cities of Russia”*[^6] was published by commercial company DaData. Both these tables contain information that is actual for the beginning of 2022, thus there may be some inaccuracies in geocoding and normalisation, but they are likely to be negligible.

[^5]: Населенные пункты России: численность населения и географические координаты. Web resource: https://data.rcsi.science/data-catalog/datasets/160/ (accessed 15.04.2024)
[^6]: Города России. Web resource: https://github.com/hflabs/city (accessed 15.04.2024)

**Methods.** A special Python command-line tool called *ru-smb-companies* was developed by the author of this paper to initially process huge and complex raw datasets from the FTS servers and transform them into a single table. The source code of this tool with a brief documentation is published in author's personal GitHub account[^7]. Here, a concise summary of the data processing workflow is provided. The data processing takes several steps.

[^7]: Russian small and medium business dataset generator. Web resource: https://github.com/PavelSyomin/russian-smb-companies (accessed 15.04.2024)

*First,* source datasets were downloaded from the FTS servers and stored on the local machine. In addition, their backup was uploaded to author's personal Yandex.Disk account, making it possible for other researchers to access the raw data even if FTS decides to remove it (as indeed happened with data on the number of employees).

*Second,* the filter-on-extract approach was used to extract necessary information from the raw archives into intermediary CSV tables. The filter-on-extract approach means that instead of unpacking the entire archive on the disk, the program reads its contents file by file without unpacking, parses the internal XML structure, extracts necessary attributes only, filters the records by activity code to leave legal companies only, and stores the result on the disk. Thus, the runtime and disk usage are substantially reduced compared to the traditional extract-then-filter method.

*Third,* the data from the intermediary CSV tables was aggregated to remove duplicated records. The reason for the duplicates is the procedure of data publication: as it was mentioned earlier, every version of the source dataset contains the full data for the time of publication rather than updates only, leading to the huge number of repeating records.

*Forth,* the addresses were normalised and geocoded to convert original administrative address format containing region, district, city and settlement name into simpler yet useful triplet of region, area and settlement, and add geographical coordinates and municipal codes (ОКТМО in Russian) for the settlements.

*Fifth,* the data on business entities was combined with the data on employees using taxpayer ID as unique identifier common to both datasets, and also converted into a panel time series table that includes yearly information for each company or individual.

<br>

# Results and discussion

# Conclusion

<br>

## References

<div id="refs"></div>