### 2024-04-17 legal-companies-geography/Paper.Rmd
First part of results and discussion, #1

---
output: 
  word_document:
    reference_doc: reference.docx
bibliography: refs.bib
#csl: style.csl
link-bibliography: False
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.id = FALSE,
  fig.width = 6.7,
  fig.height = 5,
  dpi = 300)
```

```{r imports, include=FALSE, cache=TRUE}
source("Script.R")
```

**Pavel O. Syomin[^1]**

<br>

<div custom-style="Title">

Legal Business Geography in Russia: Regional Analysis Using Tax Service Open Data

</div>

<br>

# Abstract

Text text text.

<br>

<div custom-style="Block Text">

**KEYWORDS:** w1, w2

</div>

<br>

[^1]: Perm State University, Department of Social and Economic Geography; 15 Bukireva st., Perm, 614068, Russia; e-mail: ntsp\@ya.ru

# Introduction

Geography of services is an important branch of modern economic and social geography. In Russia, it dates back to the 1960s when the pioneering research papers in this field of studies were published [@Tkachenko2016]. In foreign countries, the spatial aspects of services were treated as a part of human geography and touched in the scholarly works about retail [@Reilly1931; @Berry1967] and cities [@Brush1953] usually based on the central places theory [@Christaller1966]. A series of review research articles published in 1985–87 directly uses the term “geography of services” and cites dozens of original papers related to this topic [@Daniels1985; @Daniels1986; @Daniels1987]. Newer foreign studies in this field focus on the processes and phenomena in peripheral or smaller economies [@Hermelin2007] or on the arising and growing sectors like FinTech [@Lai2020; @Wojcik2020]. In Russia, recent research in the geography of services covers spatial properties of banking [@Shustova2015], education [@Shulgina2019], car rental [@Zyrianova2018], food service [@Preobrazhenskiy2018], service sector in general [@Lisin2020]. However, several types of services are rarely studied, e. g. business and legal ones [@Tkachenko2016].

The research on the geography of legal services in Russia is limited. To date, only an analytical review describing the legal market in Russia was published that contains a chapter on its spatial aspects [@Moiseeva2016], and yet another paper briefly addresses the role of Russia in the global legal services market without diving into the internal spatial properties of the Russian legal services itself [@Yastrebov2021]. There are also non-scientific products about Russian legal market that include its regional analysis[^2]. Despite the observed low interest to this issue from other researches, it can be argued that the geographical studies of legal services is necessary. First, legal services, as a subset of other business or producer services, play a substantial role in the development of regions and territories and in the economic growth [@Hansen1993], and sometimes act as a limiting factor, that is, the lack of professional legal advice may cause problems to entrepreneurs, especially to small ones, thus decreasing their performance [@Hitchens1997]. Next, the in-depth analysis of legal companies may give valuable insights about the business sectors they serve [@Faulconbridge2019] or increase the theoretical understanding of markets and industries in general [@Henderson2016]. Last but not least, legal services, apart from their narrow economic function, should be also treated as one of the elements of legal system in a wider sense that protect human rights and freedoms, ensure the rule of law, and act as one of the intermediaries between people, business and government that help to resolve the conflicts, increase the certainty and orderliness of social relations. This aspect of legal services is closely related to the social dimension of geography of services, and from this point of view it is not surprising that the research is also being conducted on the problem of legal services availablity [@Baxter2015].

[^2]: Анализ рынка юридических услуг в России в 2019-2023 гг, прогноз на 2024-2028 гг. Web resource: https://marketing.rbc.ru/research/41683/ (accessed 14.04.2024)

Legal services is quite a narrow and specific area of economy that does not reported separately in the official statistics, so for their analysis, other sources of information are necessary. Company-level georeferenced data is valuable for this task, because it can be easily filtered to leave the firms doing the business in the required activity area only and then aggregated by its spatial attributes to make a geographical overview. Researchers typically extract such the information from public GIS services [@Lachininsky2021] or business intelligence (BI) systems [@Moiseeva2016]. The disadvantage of this method, however, is the complexity of data preprocessing or the relatively high price of the data. Also, GIS and BI systems often do not enable access to historical data, providing real-time picture only, while the geographical studies may benefit from combining spatial and temporal properties. To overcome all these issues, an open administrative data may be used. In particular, the publicly available archives of Registry of Small and Medium-Sized Businesses (SMB registry) distributed by the Federal Tax Service (FTS) of Russia, in combination with a few other open data collections of the same governmental body, look valuable for the task in hand, and they have already been used by geographers [@Rostislav2021; @Gumenyuk2022].

The purpose of this paper is twofold. On the one hand, it aims to provide an overview of geography of legal business in Russia. The scope of the research is intentionally limited due to various organizational and technical constraints: it deals with the service providers only (the supply part of the market) and contains nothing about customers (the demand part), focuses on commercial legal companies with other market participants such as notaries, advocacy firms, non-profit legal organizations left untouched, and operates with only a portion of variables and metrics indicating the characteristics of legal business. On the other hand, the paper proposes a reproducible technique of working with FTS open data for analyzing the geography of services in Russia, particularly the services that do not reported independently from the others in the official statistics (mainly specific types of business and customer services) and thus can hardly be analyzed with the traditional economic geographers' toolkit.

<br>

# Materials and methods

**Data.** The research utilizes two primary data sources. All of them are published on the FTS website as open data.

The first dataset is called *“Registry of small and medium-sized enterprises”*[^3]. It contains full dumps of the respective registry in zipped XML files. The first dump was published in August 2016, and since then, new dumps were published every month. Each dump contains the whole registry and includes, in particular, company's names, their taxpayer IDs, activity codes according to All-Russia Classifier of Kinds of Economic Activity, address of incorporation (up to a settlement name). According to Russian laws, the Registry includes information about all commercial organizations and individual entrepreneurs that meet several criteria, the most important of which are limit on revenue (no more than 2 billion Russian rubles) and average number of employees (no more than 250). Thus, the open dumps of this registry can be treated as the most complete publicly available data source on the Russian business entities. In addition, this source has geographical reference in the form of human-readable addresses, and historical records from 2016 to now in the form of time series of monthly dumps of the registry.

[^3]: Единый реестр субъектов малого и среднего предпринимательства. Web resource: https://www.nalog.gov.ru/opendata/7707329152-rsmp/ (accessed 15.04.2024)

The second dataset is called *“Information about the average list count of employees of organisations”*[^4]. As the name suggests, it contains information about the number of company's employees. The data was initially published in 2019 and since then was updated with varied frequency, but at least yearly. As a result, the data archive includes information for the years from 2019 to 2023. For the unspecified reasons, the old versions of the dataset containing information for years from 2019 to 2021 were once deleted from the FTS website, but they had been already downloaded by the author and thus used in this research.

[^4]: Сведения о среднесписочной численности работников организации. Web resource: https://www.nalog.gov.ru/opendata/7707329152-sshr2019/ (accessed 15.04.2024)

In addition to the primary data sources, two lookup tables with the information on Russian cities and settlements are used to normalise and geocode raw addresses contained in the FTS data. The first table is called *“Settlement of Russia: population and geographical coordinates”*[^5]. It was prepared by the non-profit open data initiative called “Research data infrastructure” (RDI). The second table called *“Cities of Russia”*[^6] was published by commercial company DaData. Both these tables contain information that is actual for the beginning of 2022, thus there may be some inaccuracies in geocoding and normalisation, but they are likely to be negligible.

[^5]: Населенные пункты России: численность населения и географические координаты. Web resource: https://data.rcsi.science/data-catalog/datasets/160/ (accessed 15.04.2024)
[^6]: Города России. Web resource: https://github.com/hflabs/city (accessed 15.04.2024)

**Methods.** A special Python command-line tool called *ru-smb-companies* was developed by the author of this paper to initially process huge and complex raw datasets from the FTS servers and transform them into a single table. The source code of this tool with a brief documentation is published in author's personal GitHub account[^7]. Here, a concise summary of the data processing workflow is provided. The data processing takes several steps.

[^7]: Russian small and medium business dataset generator. Web resource: https://github.com/PavelSyomin/russian-smb-companies (accessed 15.04.2024)

*First,* source datasets were downloaded from the FTS servers and stored on the local machine. In addition, their backup was uploaded to author's personal Yandex.Disk account, making it possible for other researchers to access the raw data even if FTS decides to remove it (as indeed happened with data on the number of employees).

*Second,* the filter-on-extract approach was used to extract necessary information from the raw archives into intermediary CSV tables. The filter-on-extract approach means that instead of unpacking the entire archive on the disk, the program reads its contents file by file without unpacking, parses the internal XML structure, extracts necessary attributes only, filters the records by activity code equal to 69.10 to leave legal companies only, and stores the result on the disk. Thus, the runtime and disk usage are substantially reduced compared to the traditional extract-then-filter method.

*Third,* the data from the intermediary CSV tables was aggregated to remove duplicated records. The reason for the duplicates is the procedure of data publication: as it was mentioned earlier, every version of the source dataset contains the full data for the time of publication rather than updates only, leading to the huge number of repeating records.

*Forth,* the addresses were normalised and geocoded to convert original administrative address format containing region, district, city and settlement name into simpler yet useful triplet of region, area and settlement, and add geographical coordinates and municipal codes (ОКТМО in Russian) for the settlements.

*Fifth,* the data on business entities was combined with the data on employees using taxpayer ID as unique identifier common to both datasets, and also converted into a panel time series table that includes yearly information for each company or individual. This resulting table is used in the further analysis.

The research follows the general framework of regional analysis and uses the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. The analysis consists of a few research problems, or questions, that are listed below with the description of the relevant techniques used to find answers to them.

The first problem is the *distribution of legal firms by regions* of Russia. To answer this question, legal firms that existed in 2021 are counted by the region, and the result is visualized on the map.

The second problem is the *structure of legal business in regions* with respect to the kind of firms (either individual entrepreneurs or organisations). To answer this question, legal firms are counted by region and king, and the result is visualized on a geofacetted plot.

The third problem is the *role of legal business on the regional and national scale*, that is, either its importance within a given region or the importance of region's legal services in the Russia's legal market. To discover this issue, two calculations are made. The first one is used to better understand the role of the region in the country's legal business and is done by calculating the share of legal companies employees of a region in the total number of country's legal companies employees. The second one is aimed at estimating the relative role of legal business in region's economy and is done by dividing the number of legal companies employees in a region by the total number of employees in a region. Both two metrics are then plotted on a tile-grid map.

The fourth issue deals with the temporal dimension and reflects the *change of the number of legal firms* in the regions. Legal firms are calculated by region and year, and the results are plotted.

Several aspects are common to all the questions. First of all, despite the presence of data from 2016 to 2024, only a part from 2016 to 2021 inclusively is utilized for the analysis. The reason for this is the Russia's military operation in Ukraine and the subsequent structural transformation of the economy, international sanctions, substantial political and legal changes in the country that may have significantly impacted the legal market compared to the previous period, and even though all these changes could have been interesting to study, the purpose of the current research is to draw an overall picture of legal market in Russia in “normal situation” without any serious perturbations. Statistics for several consequent years, if not processed separately (e. g. in the case of temporal changes), is averaged using median as a function (e.g. number of legal firms for 2016–2021 in the region is the median of counts for each year from 2016 to 2021). Next, both organisations and individual entrepreneurs are accounted for. If statistical computations require the count of employees, it is assumed that an individual entrepreneur always has exactly one employee, that is, this person itself. This is because even though individual entrepreneurs in Russia can legally employ people, the corresponding information is not available in the FTS open data possibly due to personal data protection issues. Also, if an organisation has reported zero count of employees, it is still assumed that is has exactly one employee, that is, the company's owner itself.

<br>

# Results and discussion

**Average yearly count of legal firms by region** is displayed on a tile-grid colorpleth map in Fig. 1. Several conclusions can be drawn from this visualisation. First, Moscow is the only city where the number of legal firms is above 10 thousand, and this is the highest value in Russia. Next, a set of regions also has relatively high count of legal companies and individual entrepreneurs. In spatial terms, most of these regions form several groups. One of these groups is located at the south of the European part of Russia and consists of Krasnodar and Stavropol territories with Rostov, Volgograd and Voronezh regions. Tatarstan and Bashkir republics with Samara, Sverdlovsk, Chelyabinsk regions and Perm territory make up another group that may be called Ural-Volga. Some Siberian regions, namely Tyumen, Omsk, Novosibirsk, Kemerovo, with Krasnoyarsk territory, form the third group that may be named Siberian by its location. Also, Saint Petersburg city, Nizhniy Novgorod region, and Primorie territory act as single-region groups. The majority of other regions has relatively low count of legal firms, and in some of them (several republics of North Caucas and Altai and a portion of regions from the North) there are less than a hundred of such companies. The co-loocation of regions and groups with different count of legal firms suggests that there may be a sort of central-peripheries relations between them. For example, Moscow city and region, Saint Petersburg city, and Nizhniy Novgorod region are surrounded by dozens of regions with relatively small number of legal firms, and thus may act as a centers of legal business for them. The same may be true for the pairs of South and Caucasian, as well as for Siberia and Far East. It may be hypothesized that economic development of regions is positively associated with the development of legal business. Also, the population of a region may also matter.

<br>

```{r fig-1-counts-map, fig.cap="Fig. 1. Median yearly count of legal firms (both organisations and individual entrepreneurs) and shares of ogranisations by regions of Russia for 2016–21. Grey is no data"}
counts_map
```

<br>

Fig. 1 also displays the median shares of organisations among all legal firms in regions. Legal firms may be either organisations or individual entrepreneurs, so for each region, the rest part of the firms consists of individuals. The share of companies among all legal firms varies from less about 25% to almost 70%. However, a substantial part of regions has from 40% to 60% of organisations, and thus many regions may be categorized either as with a small domination of companies or as with small domination of individuals on the legal market. Several southern regions (Krasnodar and Stavropol territories, Rostov region, Kalmyk republic), as well as several northern ones (Arkhangelsk region, Komi republic, Nenetz and Yamal territories), and some central regions (Moscow and Vladimir) have substantially lower shares of companies compared to the surrounding regions, thus forming the areas of predominantly individual legal business. The overall count of legal firms seems to have to connection with the share of organisation among them.

# Conclusion

<br>

## References

<div id="refs"></div>