### 2024-04-22 legal-companies-geography/Paper.Rmd
Perform proofreading of introduction, materials and methods, add CSL style, make minor fixes in reference docx and in biblio, #1

---
output: 
  word_document:
    reference_doc: reference.docx
bibliography: refs.bib
csl: style.csl
link-bibliography: False
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.id = FALSE,
  fig.width = 6.7,
  fig.height = 5,
  dpi = 300)
```

```{r imports, include=FALSE, cache=TRUE}
source("Script.R")
```

**Pavel O. Syomin[^1]**

<br>

<div custom-style="Title">

Legal Business Geography in Russia: Regional Analysis and Mapping Using Tax Service Open Data

</div>

<br>

# Abstract

This paper provides an exploratory regional analysis of legal business in Russia for 2016–21. The analysis is based mainly on the Federal Tax Service (FTS) open data on small and medium-sized businesses and on the number of employees of organisations, as well as on portions of official statistical information. The paper proposes a reproducible technique of working with FTS open data for regional analysis in service geography that was partially implemented in a specially designed Python command-line tool. The analysis includes questions about the count of legal firms and their employees by region, about the share of organisations in regional legal business, about the role of legal business in region's economy, and about the yearly change of the number of legal companies in regions. In addition, it briefly addresses an issue of probable connection between the location of multi-regional courts and the development of legal business. It is found out that legal business in Russia is highly concentrated in Moscow and Saint Petersburg that jointly account for 1/3 of legal firms and workers and are also at the top by the role of legal business in the region. However, there are other regions with relatively high development of legal business, that are located mainly on the European South of Russia, in the Volga-Urals and in Siberia. There is a group of regions with domination of organisations in legal business and the group of regions where individual entrepreneurs prevail over companies. District courts are usually located in the regions with high level of legal business development, but the mechanism of this co-location requires further investigation. In general, the number of legal firms in regions grown from 2016 to 2021, with the common inverted U-shape pattern probably explained by the coronavirus (covid-19) pandemic that negetively impacted the small and medium business and reverted the initial growth trend.

<br>

<div custom-style="Block Text">

**KEYWORDS:** regional analysis, geography of services, legal services, legal business, open data

</div>

<br>

[^1]: Perm State University, Department of Social and Economic Geography; 15 Bukireva st., Perm, 614068, Russia; e-mail: ntsp\@ya.ru

# Introduction

Geography of services is an important branch of modern economic and social geography. In Russia, the discipline can be traced back to the 1960s, when pioneering research papers were published on this subject [@Tkachenko2016]. In foreign countries, the spatial aspects of services were treated as a part of human geography and were first touched upon in the scholarly works about retail [@Reilly1931; @Berry1967] and cities [@Brush1953], which were usually based on the central places theory [@Christaller1966]. A series of review research articles published in 1985–87 directly employs the term “geography of services” and cites numerous original papers related to this topic [@Daniels1985; @Daniels1986; @Daniels1987]. More recent foreign studies in this field have focused on the processes and phenomena occurring in peripheral or smaller economies [@Hermelin2007] or on the rising sectors such as FinTech [@Lai2020; @Wojcik2020]. Recent research in the geography of services in Russia addresses spatial properties of banking [@Shustova2015], education [@Shulgina2019], car rental [@Zyrianova2018], food service [@Preobrazhenskiy2018], and the service sector in general [@Lisin2020]. Nevertheless, several types of services remain understudied by geographers, including business and legal services [@Tkachenko2016].

The research on the geography of legal services in Russia is limited. To date, only an analytical review describing the legal market in Russia has been published, which contains a chapter on its spatial aspects [@Moiseeva2016]. Another paper briefly addresses the role of Russia in the global legal services market without delving into the internal spatial properties of the Russian legal business [@Yastrebov2021]. In addition to the scientific publications, there are business products about the legal services market in Russia that include its regional analysis[^2]. Despite a low observed interest in this issue from other researchers, it can be argued that geographical studies of legal services are essential. First, legal services, as a subset of other business or producer services, play a substantial role in the development of regions and territories and in the economic growth [@Hansen1993]. In some cases, the lack of professional legal advice may even act as a limiting factor, causing problems for entrepreneurs, especially small ones, and decreasing their performance [@Hitchens1997]. Next, a comprehensive examination of legal companies may yield valuable insights into the business sectors they serve [@Faulconbridge2019] or enhance the theoretical understanding of markets and industries in general [@Henderson2016]. Finally, legal services should be considered not only in terms of their narrow economic function, but also as an integral element of the broader legal system. This system is responsible for protecting human rights and freedoms, ensuring the rule of law, and acting as an intermediary between people, businesses and government. It helps to resolve conflicts, increase certainty and orderliness in social relations, and thus contributes to the stability and development of society. This aspect of legal services is closely related to the social dimension of the geography of services. It is therefore unsurprising that that research is being conducted on the problem of legal services availability [@Baxter2015].

[^2]: Анализ рынка юридических услуг в России в 2019-2023 гг, прогноз на 2024-2028 гг. Web resource: https://marketing.rbc.ru/research/41683/ (accessed 14.04.2024)

The field of legal services represents a narrow and specific area of the economy that is not reported separately in the official statistics. Consequently, to analyze this sector, it is necessary to draw upon other sources of information. Georeferenced data on distinct firms is of particular value in this context, as it can be readily filtered to exclude firms that are not engaged in the requisite activity and then aggregated according to its spatial attributes to make a geographical summary. Researchers typically extract such the information from public geoinformation (GIS) services [@Lachininsky2021] or business intelligence (BI) systems [@Moiseeva2016]. However, this method has two disadvantages. Firstly, the complexity of data preprocessing makes it difficult to use. Secondly, the relatively high price of the data makes it unaffordable for many researchers. Furthermore, GIS and BI systems often do not enable access to historical data, providing only a real-time picture, while geographical studies may be enhanced by combining spatial and temporal properties. To address these challenges, the use of open administrative data may be a viable solution. In particular, the publicly available archives of the Registry of Small and Medium-Sized Businesses (SMB registry) distributed by the Federal Tax Service (FTS) of Russia, in combination with a few other open data collections of the same governmental body, appear to be a valuable resource for task at hand. These data sources have already been used by geographers [@Rostislav2021; @Gumenyuk2022].

The objective of this paper is twofold. Primarily, it seeks to provide an overview of the geography of legal business in Russia. The scope of the research is intentionally limited due to various organizational and technical constraints. It focuses exclusively on service providers, or the supply side of the market, and does not address customers, the demand side. Furthermore, it concentrates on commercial legal businesses, excluding other market participants such as notaries, advocacy firms, and non-profit legal organizations. Finally, only a subset of metrics indicative of characteristics of legal businesses is included in the analysis. The specific research questions are presented in the “Materials and methods” section below.

In addition to the primary purpose, the paper proposes a reproducible methodology of working with FTS open data for analyzing the geography of services in Russia, particularly those not reported separately in official statistics, such as specific types of business and customer services. Their analysis would be difficult using the traditional tools and data sources of economic geographers.

<br>

# Materials and methods

**Data.** The research utilizes two primary data sources: the Unified Registry of Small and Medium-Sized Businesses (SMB Registry)[^3] and the Information about the Average Number of Employees of Organisations (Information about Employees)[^4]. Both of these are available on the FTS website as open data.

[^3]: Единый реестр субъектов малого и среднего предпринимательства. Web resource: https://www.nalog.gov.ru/opendata/7707329152-rsmp/ (accessed 15.04.2024)
[^4]: Сведения о среднесписочной численности работников организации. Web resource: https://www.nalog.gov.ru/opendata/7707329152-sshr2019/ (accessed 15.04.2024)

The SMB Registry dataset contains full dumps of the respective registry in zipped XML files. The initial dataset was made available in August 2016, and updates have been released on a monthly basis. Each dump contains the entire registry and includes, in particular, the company or individual names, taxpayer IDs, and activity codes according to All-Russia Classifier of Kinds of Economic Activity, as well as addresses of incorporation (up to the settlement name). According to Russian laws, the Registry includes information about all commercial organizations and individual entrepreneurs that meet several criteria (SMB subjects). The most important criteria are limits on the revenue (no more than 2B Russian rubles) and on the average number of employees (no more than 250). Consequently, the open dumps of this registry represent the most comprehensive publicly accessible data source on the Russian business entities. Moreover, the archive of these dumps constitutes a time series spanning from August 2016 to the present, thereby providing a history of small and medium-sized businesses changes.

The Information about Employees dataset contains company taxpayer IDs and average yearly numbers of employees for the previous year. The data was initially published in 2019 and has been updated with varying frequencies, though at least annually. Consequently, the data archive encompasses the period from 2018 to 2023. For reasons that are not specified, the older versions of the dataset which included information for the years from 2018 to 2022 are not currently available on the FTS website. However, the data had already been downloaded by the author prior to its removal, and thus it is used in this research.

In addition to the primary data sources, two lookup tables with the information on Russian cities and settlements are used to normalize and geocode the raw addresses contained in the FTS data. The first table is “Settlements of Russia: Population and Geographical Coordinates”[^5] prepared by the nonprofit open data initiative known as “Research Data Infrastructure” (RDI). The second table is “Cities of Russia”[^6] published by the commercial company HFLabs. Data on several cities is absent from this table, so it was entered manually. Both tables contain information that is actual for the beginning of 2022, and thus there may be some inaccuracies in geocoding and normalization. However, these are likely to be negligible.

[^5]: Населенные пункты России: численность населения и географические координаты. Web resource: https://data.rcsi.science/data-catalog/datasets/160/ (accessed 15.04.2024)
[^6]: Города России. Web resource: https://github.com/hflabs/city (accessed 15.04.2024)

Apart from these datasets, a small manually collected table with the locations of Russian district-level courts is utilized. District-level courts span their activity over two or more regions of Russia that constitute a judicial district, checking and reviewing decisions of lower-level courts. These category include appeal and cassation ordinary and commercial courts, with a total of 47 such courts in Russia. The location data was obtained from the Supreme Court of Russia website[^7] and the legal information system called “Consultant Plus”[^8]. While not the primary focus of this paper, the court location data was utilized to supplement the regional analysis by examining the relationship between the legal business geography and the placement of district-level courts.

[^7]: Информация о кассационных и апелляционных судах общей юрисдикции. Web resource: https://vsrf.ru/press_center/news/28148/ (accessed 15.04.2024)
[^8]: Справочная информация: "Арбитражные суды". Web resource: https://www.consultant.ru/document/cons_doc_LAW_56549/ (accessed 15.04.2024)

**Methods.** A Python command-line, ru-smb-companies, was developed by the author of this paper to initially process large and complex raw XML datasets from the FTS servers and transform them into a single CSV table. The source code of this tool, along with a brief documentation, is published in the author's personal GitHub account[^9]. Here, a concise summary of the data processing workflow is provided. The data processing takes several steps.

[^9]: Russian small and medium business dataset generator. Web resource: https://github.com/PavelSyomin/russian-smb-companies (accessed 15.04.2024)

Step 1 is data download. The dumps of the SMB Registry and the archives of Information about Employees were downloaded from the FTS servers and stored on the local machine. Additionally, their backup was uploaded to the author's personal Yandex.Disk account, thus enabling other researchers to access the raw data even if FTS decides to remove it (as indeed occurred with the data on employees).

Step 2 is data unpacking. A “filter-on-extract” approach was used to facilitate the retrieval of necessary information into intermediary CSV tables. This approach entails the program reading the contents of the archive file by file, rather than unpacking the entire archive on the disk. Each XML file is temporarily extracted from the archive to the memory and parsed. The program filters the records with the data on organizations and individuals contained in the XML file by activity code equal to 69.10 (“activity in the area of law” according to All-Russia Classifier of Kinds of Economic Activity). The resulting set of filtered records is then subjected to attribute extraction, and the extracted data is stored on disk. Once this process is complete, the memory occupied by the temporarily extracted XML file is freed, and the next file is being processed. The preliminary filtering by activity code decreases the runtime of the application, while the direct reading of archive contents reduces the disk space occupied by temporary files. Consequently, this step is executed relatively fast in comparison to the less optimized “extract, then filter” approach.

Step 3 is data aggregation. The intermediary CSV tables generated in the step 2 were processed to remove duplicate records. The reason for the duplicates is the procedure of data publication. As previously stated, each version of the source datasets contains the complete data for the time of publication rather than updates only. This results in a significant number of repeating records.

Step 4 is geocoding. This involves the addition of the geographical coordinates and municipal codes of the settlements where the SMB subjects are located to the original data table. Furthermore, the addresses were normalized by converting from the original administrative division format, which consists of region, district, city, and settlement names, to a simpler yet useful municipal division format with region, area, and settlement names.

Step 5 is data panelization. The data from on SMB subjects was combined with the data on employees using taxpayer ID as a unique identifier common to both datasets, and also converted into a panel time series table that includes yearly information for each company or individual. This resulting table is used in the further analysis.

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “legal firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “legal workers” refers to the employees of these firms.

Problem 1 is the distribution of legal firms across Russia's regions. To discover this distribution, the median number of legal firms in each region is calculated for the period from 2016 to 2021. This result is then visualized on a map.

Problem 2 is the structure of legal business in the regions with respect to the kind of firms (either individuals or companies). To address this issue, the median share of legal companies in each region over the 2016–2021 period is calculated and then mapped.

Problem 3 is the relative importance of regional legal business which is divided into two sub-problems. The first is the importance of a region's legal services in the Russian legal market measured by the share of legal workers in a region in the number of country's legal workers. The second is the importance of these services within a given region estimated as a share of legal workers in a region in the total number of workers in that region. Both metrics are calculated for each year from 2018 to 2021, averaged by a median function, and then plotted on a tile-grid map. Additionally, regions with district-level courts are marked on the map to analyze the relationship between the location of such courts and the development of legal business.

Problem 4 is the temporal change in the legal business sector across regions. To reach conclusions, legal firms are calculated by region and year for the 2016–2021 period. The time series of counts are visualized as line plots to identify temporal patterns. The regions are divided into the three groups based on the overall growth, decline, or stability of legal business in the observed period, and the group is also marked on the map.

<br>

# Results and discussion

**Average yearly count of legal firms by region** is displayed on a tile-grid colorpleth map in Fig. 1. Several conclusions can be drawn from this visualisation. First, Moscow is the only city where the number of legal firms is above 10 thousand, and this is the highest value in Russia. Next, a set of regions also has relatively high count of legal companies and individual entrepreneurs. In spatial terms, most of these regions form several groups. One of these groups is located at the south of the European part of Russia and consists of Krasnodar and Stavropol territories with Rostov, Volgograd and Voronezh regions. Tatarstan and Bashkir republics with Samara, Sverdlovsk, Chelyabinsk regions and Perm territory make up another group that may be called Ural-Volga. Some Siberian regions, namely Tyumen, Omsk, Novosibirsk, Kemerovo, with Krasnoyarsk territory, form the third group that may be named Siberian by its location. Also, Saint Petersburg city, Nizhniy Novgorod region, and Primorie territory act as single-region groups. The majority of other regions has relatively low count of legal firms, and in some of them (several republics of North Caucas and Altai and a portion of regions from the North) there are less than a hundred of such companies. The co-loocation of regions and groups with different count of legal firms suggests that there may be a sort of central-peripheries relations between them. For example, Moscow city and region, Saint Petersburg city, and Nizhniy Novgorod region are surrounded by dozens of regions with relatively small number of legal firms, and thus may act as a centers of legal business for them. The same may be true for the pairs of South and Caucasian, as well as for Siberia and Far East. It may be hypothesized that economic development of regions is positively associated with the development of legal business. Also, the population of a region may also matter.

<br>

```{r fig-1-counts-map, fig.cap="Fig. 1. Median yearly count of legal firms (both organisations and individual entrepreneurs) by regions of Russia for 2016–21. Grey is no data"}
counts_map
```

<br>

Fig. 2 displays the *median shares of organisations among all legal firms in regions*. Legal firms may be either organisations or individual entrepreneurs, so for each region, the rest part of the firms consists of individuals. The share of companies among all legal firms varies from less about 25% to almost 70%. However, a substantial part of regions has from 40% to 60% of organisations, and thus many regions may be categorized either as with a small domination of companies or as with small domination of individuals on the legal market. Several southern regions (Krasnodar and Stavropol territories, Rostov region, Kalmyk republic), as well as several northern ones (Arkhangelsk region, Komi republic, Nenetz and Yamal territories), and some central regions (Kaluga and Moscow) have substantially lower shares of companies compared to the surrounding regions, thus forming the areas of predominantly individual legal business. The overall count of legal firms seems to have to connection with the share of organisation among them.

<br>

```{r fig-2-shares-map, fig.cap="Fig. 2. Median yearly shares of companies among legal firms by regions of Russia for 2016–21. Grey is no data"}
shares_map
```

<br>

Fig. 3 shows the joint results of regional analysis of number of legal companies employees. Two main metrics are drawn, the first is the concentration of legal workers in the region, that is, the share of legal companies employees of a region in all country's employees of legal companies, and the second is the share of legal workers in region's labor force. The first measure roughly describes the importance of a region in country's legal business, while the second estimates the legal business' impact in regional economy. In addition, the number of multi-region courts located in a given region is shown. The overall regional distribution of legal companies employees resembles the regional distribution of legal companies (see Fig. 1) suggesting that legal companies among regions are quite similar to each other, so that there is no significant difference between simple count of companies and weighted count of companies with respect to the number of workers. The role of legal business in region's economy is distributed differently, though. It's clear that Moscow and Saint Petersburg, the two largest cities concentrating almost 1/3 of legal companies employees, are also characterized by the relatively highest portion of legal workers in total labor force. The majority of other regions is surprisingly uniform in having the small share of legal workers (0.1–0.2%) in labor force, with Sverdlovsk, Tyumen and Novosibirsk regions being rare exceptions. It is also worth noticing that regions with extremely low share in country's legal business (for example, Tyva republic and Altai territory) also usually have low share of legal workers in regional labor force, with a single exception of Evrei autonomous region. In addition, the regional role of legal business seems to have to connection with the presence or absence of multi-regional courts. However, looks like that the national role of a region in legal business is somehow connected with the location of such the courts, because almost all regions with relatively high share in country's legal companies employees (>1%) have at least one multi-regional court, and the reverse is true either: other regions typically do not have such the courts. Of course, there are regions that do not fall within this tendencies (Bashkortostan republic and Volgograd region, as well as three sub-Moscow regions, Vologda region and Zabaikal territory), but these deviations may be explained by various factors (e. g. for Tula, Vladimir and Kaluga regions the presence of multi-regional courts is not associated with the development of legal business probably due to the proximity to Moscow, so that the need in local legal firms does not arises).

<br>

```{r fig-3-employees-map, fig.cap="Fig 3. National and regional median yearly shares of legal companies employees in regions of Russia in 2018–2021"}
employees_map
```

<br>

The changes in the number of legal firms by region are depicted in Fig. 4. It can be clearly seen that the count of such companies increased in the majority of regions, indicating the development of legal business. Few regions have lost companies, all of them are located in the European part of Russia, and their net change in the number of legal firms from 2016 to 2021 is relatively low (usually less than 10%) compared to the increase rates in other regions. Sometimes neither decline or growth was not clearly expressed (for example, Altai republic, Lipetsk and Tambov regions). The regions with the largest growth are some republics at Notrh Caucas, as well as Kalmyk republic and Nenetz territory. Their results are probably explained by the low base effect, because the absolute counts of legal firms in all there regions are below 100 (see Fig. 1). As for other regions with high growth, Leningrad and Moscow regions should be remarked, because their results may be partially explained by the relocation of companies from the cities (Saint Petersburg and Moscow) to the surrounding regions. However, Moscow and Saint Petersburg also gained legal firms with one of the top net changes. The observed patterns of changes indicated by lines of the map can be divided into several groups. The first is the stable yet saturating increase typical for Moscow and Saint Petersburg regions, and, less expresses, Khabarovsk territory, Rostov and Ryazan regions. The second is the inverted U-shape curve which is charasteristic of the majority of regions with overall growth in the number of legal firms. This pattern can be explained by the coronavirus (COVID-19) pandemic effect that has negatively impacted small and medium-sized business and reverted the previous trend of the growth. The “height” of this inverted U varies by region (the highest observed in Chukotka territory, Tyva and Adygeya republics). It can be hypothesised that the first group is in fact a subset of the second, but the intrinsic growth rate before the pandemic was higher, resulting in the slowing down of the growth rather than in the final decrease. The third pattern is typical for the regions with overall decrease in the number of legal firms, and it can be described as “initial short-term growth, then continuous decline”. The fourth group is a mixture of increase and decrease, for example Nenetz territory and Northern Osetia republic.

<br>

```{r fig-4-timeline-map, fig.cap="Fig. 4. The change in the number of legal firms by region in 2016–2021"}
timeline_map
```

<br>

# Conclusion

The conducted regional analysis discovered several spatial features of legal business in Russia. First, the regions are quite different with respect to the count of legal firms or their employees. Several regional clusters with relatively high concentration of legal business can be found out, for example, the European South, Volga-Ural and Siberian. Unsurprisingly, Moscow and Saint Petersburg cities are characterized by the top count of legal firms and workers, but the overall concentration is really huge, because they jointly account for approximately 1/3 of Russia's legal businesses and legal companies employees. Besides, it's again Moscow and Saint Petersburg that have the highest share of legal companies employees in region's labor force (above 0.4% that is 2 or 3 times higher than the average value for other regions). However, there are some other regions with relatively high role of legal business in regional economy: Sverdlosvk, Tyumen and Novosibirsk.

Regions can be roughly grouped into two categories: one with the legal organisations prevailing in the legal business, and the other with individual entrepreneurs prevailing over companies. The high role of companies is typical for Moscow, Saint Petersburg, some Volga and Siberian regions, as well as a small portion of central regions (Vologda and Ivanovo), while individuals are important in the European South and North, in the majority of Central European regions, in several regions of the Urals, and in the Far East.

The role of region in country's legal business is usually associated with the presence of one or many multi-regional courts in the administrative center of this region. Only two regions do not act as a host for any of such the courts but still have relatively high level of development of legal business: Volgograd region and Bashkir republic. This observation may be explained in two ways: either the courts act as a magnets for legal business, because courts are necessary for many legal activities, or the courts are placed in regions and cities with highly developer legal business. Further research may throw the light on the direction of this court-business relation.

From 2016 to 2021, the number of legal firms in Russia's regions mostly increased. The decrease is characteristic only of few regions. The top growth was observed in Moscow, Saint Petersburg, as well as in the surrounding regions. In addition, some regions (mainly at the European South) also shown very high growth, but probably due to the low base effect. The yearly patterns of change may be categorized into four groups, and the inverted U-shape change explained likely by the coronavirus (covid-19) pandemic is the most common among the regions.

The results of this explanatory study provide an overview of Russia's recent legal business geography and may be used as a support for data-driven business and governmental decisions. Several research hypothesis also may be conducted from this analysis, thus this study may be treated as a base for further academic development in the field of legal services geography in Russia. Apart from this, the reproducible data-driven technique of this research based entirely on open data and open source software tools may be used (with minor changes) for regional analysis and mapping of other narrow service sector areas.

<br>

## References

<div id="refs"></div>