### 2024-05-14 legal-companies-concentration-and-availability/journal-article/Article.Rmd
Modify plots, edit results and discussion, #2

---
output:
  word_document:
    reference: reference.docx
bibliography: refs.bib
csl: style.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.id = FALSE,
  fig.width = 6.6,
  fig.height = 5)
```

```{r imports, include=FALSE, cache = TRUE}
library(here)
source(here("journal-article", "Script.R"))
```

IRSTI 06.61.53

<div custom-style="Title">

Provision of Commercial Legal Services in Russian Cities

</div>

<div custom-style="Author">

Pavel O. Syomin

<br>

Perm State University, Russia, Perm

<br>

e-mail: ntsp\@ya.ru

</div>

<div custom-style="Abstract">

**Abstract:** The article examines the relationship between the concentration of legal companies and the availability of legal services. To conduct the analysis, a dataset was created using open data from the Federal Tax Service of Russia. The dataset includes information on the location, revenue, expenditure, and number of employees of small and medium-sized legal companies. The concentration is calculated by dividing the number of employees of legal companies in the city by the total number of employees of all legal companies in Russia. Availability is measured as the number of employees of legal companies in the city per 100,000 population. Data visualization and linear regression were utilized to estimate the hypothesized relationship. The study showed a positive linear relationship between concentration and availability. However, there are multiple approaches to achieving good availability of legal services.

**Key words:** legal services, law firms, availability, cities, market saturation, small and medium-sized businesses, open data

</div>

<div custom-style="Title">

Ресей қалаларында коммерциялық заң қызметтерін көрсету

</div>

<div custom-style="Author">

Павел Олегович Сёмин

<br>

Пермь мемлекеттік ұлттық зерттеу университеті, Ресей, Пермь

<br>

e-mail: ntsp\@ya.ru

</div>

<div custom-style="Abstract">

**Аннотация:** The article examines the relationship between the concentration of legal companies and the availability of legal services. To conduct the analysis, a dataset was created using open data from the Federal Tax Service of Russia. The dataset includes information on the location, revenue, expenditure, and number of employees of small and medium-sized legal companies. The concentration is calculated by dividing the number of employees of legal companies in the city by the total number of employees of all legal companies in Russia. Availability is measured as the number of employees of legal companies in the city per 100,000 population. Data visualization and linear regression were utilized to estimate the hypothesized relationship. The study showed a positive linear relationship between concentration and availability. However, there are multiple approaches to achieving good availability of legal services.

**Кілт сөздер:** legal services, concentration, availability, small and medium-sized businesses, open data

</div>

<div custom-style="Title">

Обеспеченность коммерческими юридическими услугами в городах России

</div>

<div custom-style="Author">

Павел Олегович Сёмин

<br>

Пермский государственный национальный исследовательский университет, Россия, Пермь

<br>

e-mail: ntsp\@ya.ru

</div>

<div custom-style="Abstract">

**Аннотация:** The article examines the relationship between the concentration of legal companies and the availability of legal services. To conduct the analysis, a dataset was created using open data from the Federal Tax Service of Russia. The dataset includes information on the location, revenue, expenditure, and number of employees of small and medium-sized legal companies. The concentration is calculated by dividing the number of employees of legal companies in the city by the total number of employees of all legal companies in Russia. Availability is measured as the number of employees of legal companies in the city per 100,000 population. Data visualization and linear regression were utilized to estimate the hypothesized relationship. The study showed a positive linear relationship between concentration and availability. However, there are multiple approaches to achieving good availability of legal services.

**Ключевые слова:** юридические услуги, юридические фирмы, обеспеченность, города, насыщенность рынка, малое и среднее предпринимательство, открытые данные

</div>

# Introduction

Geography of services is an important and well-established field of geographical studies. Its history dates back to the second half of 20th century [@Tkachenko2016]. Two research approaches exist in geography of services: the first can be described as economic, and the second can be designated as social. The economic approach, as its name suggests, treats the services as a regular part of the economy, dealing with their role in GDP or employment, their impact on the growth, the interplay between services and other business activities. For example, a research by D. Wójcik [-@Wojcik2020] provides an overview of FihTech geography in an economic-based manner. The social approach, while closely related to the economic one, focuses on the services as the foundation of people's well-being and studies services provision, availability and accessibility, as well as the diversity and quaility of services. For instance, the study in Indonesia [@Pratiwi2021] uses the various socio-economic data from the national health insurance system grouped by ~500 country's districts to analyze the relationship between a set of public health quality metrics and discuss, among the other problems, the equity problems arising from the geographical differences. The issue of access to various services has been of great importance for decades [@Blacksell1990; @Keleher1996] because it is directly connected to the quality of life and ensuring of human rights. Research has also been conducted on the links between market properties and service availability [@Lu2020].

Cities have traditionally been viewed as regional centers that provide goods and services to smaller surrounding settlements [@Christaller1966]. The urban studies are thus tightly connected with the geography of services [@Reilly1931; @Brush1953]. It is well-known that professional and business services tend to cluster in large cities, and their spatial distribution is closely linked to urban hierarchies [@Keeble2002; @Rubalcaba2013; @Kekezi2019]. The concentration of businesses provides economic benefits for the companies and thus is likely to become self-sustaining. The common sense suggests that such concentration does not only leads to the economic growth, but also positively impacts the service provision, thus being productive in both economic and social dimensions. It is usually the largest cities or the regional centres where the local communities benefit from the variety and quantity of diverse service providers, compared to smaller settlements and rural areas, where the availability of services is limited due to the low number of their providers. However, is this relationship true indeed? The common sense can easily turn out to be wrong, and thus the rigorous data-driven testing of the connection between the number of service providers and the provision of services is crucial.

This paper empirically examines the relationship between the number of law firms in Russian cities and the provision of legal services. Typically, the literature on services availability focuses on medical services [@Joseph1982], but seems that legal services, though being understudied by geographers, are also worth attention [@Blacksell1988; @Patel2008], because they are directly related to the rule of law, protection of human rights, supporting orderliness and stability of social relations. The analysis is based on the open data about small and medium-sized businesses (SMB) published by the Federal Tax Service (FTS) of Russia. The main research question is how the specific availability of legal professionals is affected by the concentration of law firms. Additionally, the study aims to find out geographical patterns in the spatial distribution of legal services availability.

# Materials and methods

**Data.** This research utilizes a geo-referenced dataset on small and medium-sized law firms incorporated in Russia. The dataset is created from FTS open data dumps (archives) of the state SMB Registry[^1] and FTS open data on the average number of employees of ogranizations[^2] using the Python command-line tool *ru-smb-companies*[^3] developed by the author of the paper. It contains information about the particular companies, including company's name, taxpayer ID, region and settlement of incorporation, main activity code. Law firms are selected by the main activity code equal to 69.10, and all the other companies were dropped from the source data. This technique of selection by the main activity code was derived from the analytical paper on the Russian legal services market prepared by the members of the Institute for the Rule of Law [@Moiseeva2016]. Additionally, the dataset was filtered by the following conditions: the year is 2021, the company is an organization rather than a sole entrepreneur, the company has declared non-zero revenue and expenditure for the target year (implying that the company is active), and the company's address of registration is in a city rather than in a rural area. In addition to the main dataset, a data on cities population freely distributed by the commercial company HFLabs is used [^4].

[^1]: https://www.nalog.gov.ru/opendata/7707329152-rsmp/
[^2]: https://www.nalog.gov.ru/opendata/7707329152-sshr2019/
[^3]: https://github.com/PavelSyomin/russian-smb-companies
[^4]: https://github.com/hflabs/city

During the exploratory data analysis, six cities are identified as outliers subsequently excluded from the data. Two of these cities, Moscow and Saint Petersburg, are excluded due to the extremely high number of law firms employees: almost 18,866 and 5,550 respectively, while the next city has only less than 2,000 employees. The four other cities (Innopolis, Kirovsk, Bronnytsy, Krasnoarmeisk) are characterized by the very high proportion of law firms employees in city's population (from 38 per 10,000 in Krasnoarmeisk to 417 in Innopolis) compared to less than 20 per 10,000 in all the other cities.

**Methods.** A linear regression is the main method of data analysis used in this study. An independent variable for the regression is the number of employees of law firms in a city denoted as *n*. Theoretically, both number of law firms or their employees may be used, but law firms in different cities may have different size, thus the number of employees seems to be more stable and consistent metric. The value for each city is calculated as a sum of numbers of employees of each law firm incorporated in this city. If a law firm reported zero count of employees, it is assumed that the number of employees is equal to one. The dependent variable in the regression is specific provision of legal services which is denoted as *p* and calculated by dividing the number of employees of law firms in the city by the city's population and multiplying the result by 100,000. In other words, it's a number of law firms employees per 100,000 of population. This formula is a modified version of the legal market saturation metric used in the analysis of the Russian legal services market conducted by the Institute for the Rule of Law [@Moiseeva2016]. In addition, a grouping variable *g* categorizing the cities by population is used to check whether the overall relationship is dependent on the city size. The grouping by population is performed on the basis of the classification system of G. M. Lappo [-@Lappo1997]. Two linear models are fitted, one without grouping and one with grouping. The results are visialized on a plot. The analysis was performed using the R programming language and additional packages. The source code for this paper can be found in the public repository[^2].

[^2]: https://github.com/PavelSyomin/ru-smb-companies-papers/tree/main/legal-companies-concentration-and-availability

**Theoretical model.** Despite the perceived simplicity of the modelling described earlier, it has a theoretical mathematical background, and several possible types of outcome may be laid out right now before the analysis of the data in hand. It is known that a linear model is described by the two coefficients, intercept (a free term of the linear equation showing the relative shift of the regression line along the y axis) and slope (a coefficient before the independent variable showing the angle between the regression line and x axis). The slope can be either positive, negative, or zero. As a special edge case, it can be also infinite (i. e. regression line is vertical), even though, strictly speaking, such a dependency is not a function. Thus, the linear fit described earlier can yield a result of one of four types, each corresponding to one of the slope options and, subsequently, the pattern of relationship between the number of law firms and provision of legal services. These basic theoretical types are shown in Figure 1.

```{r figure-1-theoretical-models, fig.cap="Figure 1 — Four plots showing four types of the theoretical relationship between the number of law firms employees and provision of lwgal services in cities"}
theoretical_models_plot
```

Type 1 corresponds to the zero slope. In this case, the regression line is horizontal. From the substantial point of view, it describes the absolute equality in the provision of legal services among cities, that is, the provision is independent of the number of law firms employees. Despite being quite unlikely, this type also may be described as optimal, because it guarantees equitable access to legal services (at least in spatial terms). Probably this type requires some sort of governmental intervention to form (e. g. encouraging the establishment of law firms in small cities).

Type 2 is characterized by the positive slope and, consequently, positive linear relationship between the number of law firms and the provision of legal services. It seems that at the free market this type is the most probable, even though the specific value of the slope may vary. In such case, the bigger the city, the easier it is to find a lawyer, and vise versa.

Type 3 is the reverse of Type 2: the slope of the regression line is negative, and so is the relationship between the number of law firms and the provision of legal services. It is also quite improbable. One possible situation in which this type may form is the unexpected and rapid migration of people from the smallest cities to the largest with almost uniform (Type 1) relationship as a prerequisite. Under such circumstances, the relative number of law firms in small cities increases fast, but the relative number of law firms in large cities can not grow immediately as fast as the population, and the largest cities start to suffer from the lack of legal services, while the smallest cities are oversaturated with law firms.

Type 4 is purely theoretical, but important in conjunction with the three other types. As Type 1, it also signals about the independence between the number of law firms and provision of legal services, but here the number of law firms employees is fixed rather than their proportion in the population. Again, it is very unlikely to happen in the real world, but it also suggests that some sort of governmental intervention (e. g. strictly controlling the number of law firms) is taking place.

# Results and discussion

Figure 1 shows the relationship between the concentration of legal companies and the availability of legal services, as measured by the metrics proposed in the Materials and methods section. The overall connection between the number of law firms and provision of legal services appears to be nonlinear and corresponds to a square root formula. However, the same relationship within each city size group is fitted well with a linear models. The coefficients of these models are different, and the slope decreases when the city population increases. While the regression lines for the small and medium cities are close to Type 4 (provision is independent of the number of lawyers because of the small variation in the number of lawers), the modelled relationship in millionaire cities resembles Type 1 (another option for independence, but due to relatively uniform and equal provision regardless the number of lawyers). The cities from big to extra-large are characterized by the positive relationship between the number of lawyers and the provision of legal services and can be classified as Type 2.

<br>

```{r fig-1-concentration-and-accessibility, fig.cap="Figure 1. The relationship between the concentration of legal companies and the availability of legal services. City population group and relative city rank within the region are marked. The formulas used to calculate the metrics are described in the Methods section. Both the X and Y axes are log10-scaled. Dotted regression lines represent the linear fits within city population groups, while the solid gray line represents the overall linear fit. The regression line for the millionaire cities is fitted excluding the outliers (Moscow and Saint Petersburg)", fig.height=4}
ca_plot
```

In general, regional centers have relatively high number of lawyers and considerable level of legal services provision. This observation is probably explained by the centralization tendencies in the regions and the concentration of business and services in their administrative centers that situate the majority of government bodies and are also usually the most populous cities within a region.

Moscow and Saint Petersburg, though excluded from the data analysis and not shown on the main figure, are two remarkable outliers with extremely high concentration of legal services and a good level of legal services availability. Moscow alone accounts for almost one-third of all legal companies' employees, and Saint Petersburg accounts for about 10%. The position of the corresponding data points relative to the main regression lines suggests that there is an “availability saturation” level, where further concentration has little effect on availability.

The general tendencies observed in the Figure 1 should be supplemented by the analysis of the source data distribution shown in Figure 2 that extends the Figure 1 and displays the binned count of cities according to the number of lawyers and the provision of legal services. The overwhelming majority of cities is characterized by low level of both legal services provision and law firms development. Only a limited number (18) of cities is characterized by the high (more than 500) number of lawyers. It should be noted that high number of lawyers is good for legal services provision because all these cities are at least at the medium level of provision, and none of them has less than 5 lawyers per 10,000 people. However, cities with relatively low number of laywers can nevertheless reach top performance from the point of legal services provision that is on par with the cities with the highest number of lawyers. Moreover, it appears that such small yet well-served cities are even more typical than cities with high level of both lawyers and provision. The relative location of bin counts in the Figure 2 implies that there — besides the dominant bottom left corner — there are two main clusters of cities showing two different approaches of reaching high level of legal services provision. One approach is increase both in absolute number of law firms and relative number of lawyers per population probably charasteristic of the largest cities. Another is a relatively small yet sufficient number of lawyers typical probably for small or medium cities. Thus, apart from the general tendencies, the two-ways structure is found out.

Figure 2 displays the spatial distribution of cities based on the availability of legal services. Regional centers exhibit a high relative level of availability. The most prominent cluster of good access to legal services is formed by Moscow and its satellite cities. Other notable centers with high availability include Krasnodar and Stavropol in the southern part of Russia's West, Yekaterinburg and Chelyabinsk in the Urals, and Tuymen and Novosibirsk in Western Siberia. There are regions with limited access to legal services, such as the bordering regions in the south of Siberia, as well as the Komi and Kalmyk Republics.

<br>

```{r fig-2-map, fig.cap="Figure 2. Legal services availability in Russian cities on the country's map", fig.height=4}
spatial_plot
```

# Conclusion

The concentration of legal companies in cities has a positive linear relationship with the availability of legal services. This trend is observed in all the city population groups. The largest cities in regions typically have relatively high availability of legal services, as do cities with population above 1 million. Moscow and Saint Petersburg are characterized by the highest concentration and high availability, with Moscow alone accounting for almost one-third of all legal companies employees in the country. Empirical data suggests that there may be a saturation limit beyond which the increase in concentration does not have a substantial impact on availability. There are multiple paths to achieving high levels of legal services availability, with the increasing concentration of legal companies in a city being just one option. Even cities with relatively low concentration can achieve similar levels of availability as larger ones. Moscow and its region are a cluster of areas with high legal services availability. Similarly, Saint Petersburg, Krasnodar, Stavropol, Yekaterinburg, Chelyabinsk, Tyumen, and Novosibirsk also have good access to legal services.

## References

<div id="refs"></div>