### 2025-01-27 law-firms-names/Paper.Rmd
Revert from umap-dbscan to kmeans, reassign strategies and categories, switch from clusters to categories, make a dendrogram of categories with strategies and a map of localmoran i, make other changes, delete useless code from script, #4

---
title: "Identification and Spatial Analysis of Law Firms Naming Patterns in Russia Using Text Embeddings"
author: "Pavel Syomin"
output:
  word_document:
    reference_doc: "journal-article/reference.docx"
bibliography: "journal-article/refs.bib"
csl: "journal-article/apa.csl"
link-bibliography: False  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 4.96, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples
```

```{r package-imports, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(here)
library(knitr)
library(readr)
library(readxl)
library(sf)
library(stringr)
library(tidyr)
library(units)
```

```{r data-imports, cache=TRUE, include=FALSE}
source(here("journal-article", "Script.R"))
```

<div custom-style="Affiliation">

Perm State University, Geographical faculty, Department of social and economic geography, Perm, Russia

E-mail: ntsp\@ya.ru

ORCID: 0000-0002-4015-9206

</div>

<div custom-style="Abstract">

**Abstract:** The present study analyzes law firm naming in Russia, with a particular focus on the geographical aspects of this phenomenon. The primary objective of this research is to ascertain the impact of the spatial relations between the country's regions, specifically those of a neighboring, proximity, and belonging to a parent-level subdivision, on the naming of law firms. To conduct the analysis, data regarding law firms is extracted from the open dumps of the state registry of small and medium-sized businesses. The names of the firms are transformed into numerical vectors, subsequently clustered with the KMeans method, and then manually grouped to identify naming strategies. Numerical representations of law firm names within each region are averaged to get region embeddings, which are then used to compare the names of law firms in different regions. The research findings indicate that regions situated closer to one another or sharing a common border tend to adopt slightly more similar law firm names. Furthermore, regions that are part of the same economic region also display a greater degree of similarity in law firm naming. Three primary strategies of law firm naming are “Form”, “Law“, and “Service”. Volga and North Caucasus economic regions are characterized by the predominance of the “Service” strategy, whereas the Volga-Vyatka economic region exhibits a “Law” strategy, and the remaining territories demonstrate a combination of naming strategies. The results of the study lend support to the widely used classification of economic regions in Russia. The findings also underscore the impact of neighboring, proximity, and interconnectivity on the economic and cultural attributes of the regions. The administrative open data processing techniques, as well as the methods of working with firm names based on neural networks proposed in this paper, may be used by other economic geographers for their own research purposes.

</div>

<div custom-style="Keywords">

**Keywords:** firm names; law firms; text embeddings; economic regions, text similarity; large language models.

</div>

# Statements and Declarations

**Competing Interests.** No funding was received for conducting this study. The author has no financial or proprietary interests in any material discussed in this article.

\newpage

# 1. Introduction

The geography of services, also known as the geography of the service sector, is a crucial component of geographical studies in Russia. This field has a history spanning several decades, with origins dating back to the 1960s. The geography of services has two distinct dimensions: economic and social. The former is primarily concerned with the business aspects of the services, while the latter addresses the implications for human well-being arising from the geographical organization of services [@Tkachenko2016]. The existing scientific literature on the geography of services in Russia covers a broad range of topics, including education [@Shulgina2019], car rental [@Zyrianova2018], catering [@Preobrazhenskiy2018], medical care [@Kazantcev2022; @Panin2023; @Nikoloski2024], and internet access [@Nagirnaya2015; @Zemtsov2022]. However, some areas remain underexplored, with legal services being one such example [@Tkachenko2016; @Moiseeva2016].

Despite the lack of interest from geographers, the spatial aspects of the activity of law firms are worth investigating for a number of reasons. Firstly, legal services can be categorized as business or producer services, and this category of services is important for economic development, on both a national and regional scale [@Hansen1993; @Faulconbridge2019]. Secondly, despite their comparatively minor direct economic impact, these services play a significant role as intermediaries and facilitators, contributing to the establishment and maintenance of the rule of law, social order, stability, and predictability of social relations. Thirdly, for the general public, access to legal aid is vital for the protection of human rights and the attainment of social justice. In this regard, the spatial aspects of law services are on par with the spatial issues of medical care [@Blacksell1988; @Blacksell1990].

The study of firm names — which are sometimes also referred to as company names, brand names, or simply “brands” — is a relatively uncommon yet potentially fruitful approach to the geographical analysis of services and related problems. A firm name is designed to distinguish the business from competitors, to highlight the sector or geographical area in which it operates, and to convey information about the services offered by a provider and the associated benefits. A brand is therefore capable of transferring a substantial amount of information in a brief textual format. Such information may prove useful in uncovering the city's business landscape and consumer perceptions of a particular city with its distinctive services [@Raento2001]. It may also help to analyze trends common to a specific sector in a geographic region [@Marjavaara2022]. The incorporation of place names (toponyms) in brands influences the business development and consumer attitudes [@Peluso2012; @Zhang2014; @Erdogan2016; @Noorda2019]. Company names encapsulate a multitude of intertwined properties relevant to the economy and culture of a given location. This synthesis makes them suitable for the identification of regions, especially in the case of vernacular regions [@Reed1976; @Liesch2015; @Vukosav2016; @Holtkamp2018].

Law firm names may follow a number of patterns. A pattern can be defined as as a typical model, or pattern, of brand construction identified by usage of identical or semantically similar words or phrases. For example, names with the words “law firm” or “law company” follow one pattern, and names with the words “legal aid” are an example of another pattern. Patterns may be grouped multiple times forming a tree-like classification [@Anora2015]. Such classification may become a basis for identifying naming strategies. Compared to naming patterns, naming strategies are broader groups of firm names that are usually based on sharing a common purpose — e.g. emphasis the belonging of a company to a certain activity type. @BassolsiGardella2023 argue for a two-dimensional Strategic Naming Model (SNM) that includes two pairs of strategies: belonging and uniqueness, no-strategy and market. Pairs may be represented as X and Y axes of a plot, respectively, with paired strategies taking opposite positions on these axes: belonging is an opponent of uniqueness, market is an opponent for no-strategy. Authors of the aforementioned paper discover a difference in naming strategies composition of tourist industry firms in three Columbia cities. Probably such a variation is not be limited to Columbia cities and tourism and can be found in the case with law firms in Russia.

The aim of this research is twofold. First, it tries to identify law firms naming patterns in Russia and categorize them following the framework of SNM. Second, it seeks to check whether there are regularities in the spatial distribution of law firms naming patterns. The potential viability of these two statements cannot be directly conducted from the existing research, but it relies on a common-sense observations about the similarities and differences of law firm names in and between various regions of the country. According to the famous Tobler's First Law of Geography, “near things are more related than distant things” [@Tobler1970], so there is a chance that law firms naming patterns may form spatial clusters and hotspots that at least a bit reflect some economic and, to a lesser extent, cultural similarities of the territories where these firms are located. Given the vast area and a material heterogeneity of economical and cultural landscape of Russia, these names should not have a uniform spatial distribution. However, it is just a hypothesis that needs testing.

# 2. Materials and methods

## 2.1. Data

The issue of obtaining information about firm names is a challenging one. All Russian organizations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data resell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized enterprises (SME registry) may be used [@SMERegistry2024]. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. As a rule of thumb, all companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SMB registry, without any action being required on their part. There are additional qualifications for the SME registry, e. g. absence of non-SME or foreign control such as, but these should not affect a material portion of law firms. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SMB registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organizations and individuals.

The downloadable version of the registry includes, in particular, firm names, addresses of registration (up to a settlement), and codes of activity kinds compatible with NACE Rev.2. It should be noted that the registry does not include several types of firms, e.g. foreign commercial law firms without a juridical person registered according to Russian law, non-profit organizations, bar (advocacy) establishments, large companies etc. This results in some data being absent. Nevertheless, this issue should be considered insignificant for a variety of reasons. Foreign law firms usually register their affiliated juridical person in Russia and thus are not omitted in the registry. Non-profit organisations' share on Russian legal market does not exceed 1% [@Moiseeva2016]. The names of bar establishments are usually quite uniform, containing the kind of the establishment (cabinet of an advocate, bureau of advocates, college of advocates, juridical consultancy) accompanied with a personal name(s) of advocates or with a mention of geographical location of a particular establishment, thus their potential spatial heterogeneity is not characteristic. A matching of the list of the law firms obtained from the SME registry and active in 2023 with the extract of law firms from the Russian Financial Statements Database for the same year shows that ~3,300 of the firms present in SME registry are absent from RFSD, and ~5,200 firms present in RFSD are missing in SME registry. The balance of missing records is not that far from the equality, and with the total number of law firms in SME registry (~34000) the share of missing entries is below 15%. Consequently, the completeness of information is relatively high, and it can be used to draw reliable conclusions.

## 2.2. Methods

### 2.2.1. Law firms selection criteria

In this study, the term “law firm” is defined as a commercial organization registered in Russia in accordance with the Civil Code of Russia, with a NACE Rev.2-compatible main activity code equal to 69.10 (activity in the area of law). The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, advocacy firms, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SME registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

The data used in the study covers a period of six years, from 2016 to 2021. The initial point in time is defined by the commencement of SME registry data availability. The finish point, 2021, is the latest year, is the latest year with “normal” conditions in Russia not impacted by military operation in Ukraine and consequent international sanctions and domestic economic and political perturbations. The purpose of temporal dimension is to exclude short-living law firms that are likely to be dummy companies. Organizations with a lifespan of less than three years have been omitted from the research, while all the remaining have been included in the further analysis. In addition, the utilization of multi-year data acts as a form of averaging, smoothing the potential distortions of spatial structure resulting from the random year-to-year changes of active law firms list.

### 2.2.2. Firm data acquisition

The processing of SME registry data is conducted in a series of stages. In the initial phase of the research process, all source archives are downloaded. Each source archive contains a full copy of the registry data for the specified publication date, with a new archive becoming available on a monthly basis. Subsequently, data pertaining to law firms is extracted from the archives and stored in CSV files. Each CSV file contains data from one source archive. Subsequently, all CSV files are consolidated into a single table, the duplicated records are removed, and the remaining records are augmented with start and end temporal markers. Subsequently, the addresses are normalized to resolve the inconsistencies in location names and geocoded up to a city/settlement level to add geographical coordinates. The lack of within-city location in the source data prevents from more fine-grained geocoding but does not pose a serious problem because city-level coordinates are enough for large-scale analysis on country or region level. Finally, the resulting table is transformed into a panel view, with each row representing a single law firm at a given year. All these stages are implemented in an open-source Python command-line tool.

### 2.2.3. Firm name embeddings retrieval

The Yandex Foundation Models cloud service [@Yandex2024] is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organizational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed. The dimension of a firm name vector is 256. Contextual embeddings (YandexGPT) were used in favor of static ones (e.g. FastText), because during the preliminary analysis, they have resulted in far better clusterisation of law firm names (see the description of this step below) which is crucial for naming strategies identification.

### 2.2.4. Naming strategies identification

The dimension of law firm names vectors is reduced from 256 to 2 with the UMAP method, and the resulting 2D vectors are clustered using the DBSCAN technique and plotted to illustrate the results. The clusters are subsequently subjected to manual analysis. A random sample of 20 names is selected from each cluster. The names samples are carefully examined, and each cluster is attached a concise description (e.g. “Consulting companies”), a category, and a strategy. Clusters with descriptions, categories and strategies form a tree-like hierarchy where clusters make up categories, and categories belong to strategies.

The concept of naming strategies is derived from the works of @Arora2015 and @BassolsiGardella2023. The latter paper proposes a strategic naming model with four broad naming strategies, namely “Uniqueness”, “Belonging”, “Market”, and “No-strategy”. Thee strategies form a 2×2 matrix, where two pairs of “Uniqueness vs Belonging” and “Market vs No-strategy” are axes, and strategies in the pairs are opposite to each other. Though invented for classification of names of companies in tourism, this naming framework fits well for the purpose of law firm names grouping. The specific descriptions and their mapping to the categories and strategies are provided, explained and analysed in the results section below.

### 2.2.5. Spatial regularities identification

Analysis of autocorrelation is used to check whether the spatial distribution of names belonging to each cluster identified on the previous step is random or has some regularities. Regions (constituent entries) or the Russian Federation were selected as a spatial subdivision for calculating the autocorrelation measures. It could have been more reasonable to use the next spatial level, municipal territories, to perform the computations, but exploratory data analysis has shown that a lion's share of such territories has no law firms, because the legal business is highly concentrated in cities and large settlements.

Six spatial weights matrices are used to ensure the robustness of the results. Three of these matrices are constructed with K nearest neighbours technique based on the location of region's centroids, where K equals 3, 4 or 6, respectively. Another matrix also uses geographical properties of regions, but instead of centroids, it relies on the existence of common border lines or points to identify neighbouring regions. Sakhalin and Kaliningrad regions, being enclaves and thus having zero immediate neighbors, were manually connected with a few closest regions. Two remaining matrices treat regions belonging to the same upper-level metaregion as neighbours, and all the other regions as non-neighbours. For one of these two matrices, economic regions act as metaregions, and for another, federal districts are used in this fashion. Economic regions are defined according to the classification widely used by economic geographers and spatial development practitioners in Russia for decades and outlined, for example, in [@Leyzerovich2010]. Federal districts composition is defined in the Executive Order of the President of Russia No. 849, Dated May 13, 2000, “On the Plenipotentiary Representative of the President of the Russian Federation in the Federal District” (with up-to-date amendments).

For each cluster, a global Moran's I metric is calculated. The variable of interest is a share of law firm names belonging to the cluster in a total quantity of law firm names in a region. If there are no law firm names of the cluster in the region, the share is considered to equal zero. P-values are adjusted for multiple comparisons with Holm's correction. Significance level (alpha) is set to 0.05.

For the clusters with statistically significant global Moran's I, a local Moran's I was subsequently calculated. P-values are adjusted with local Holm's correction. Cluster classification of local Moran's I are mapped.

### 2.2.6. Data statistics

There are `r initial_firms_count` law firms in the SMB registry in the selected period with `r nrow(unique_names)` unique firm names. From the initial dataset, `r valid_firms_count` law firms have existed for at least three years and have been selected for the analysis. Their names sometimes are duplicated, and the number of unique law firm names being analysed is `r`.

# 3. Results and discussion

## 3.1. Law firms naming strategies

Law firm names clusters resulted from UMAP → DBSCAN pipeline are illustrated using scatterplot in Fig. 1. The visual quality of clusterisation is high: all the clusters are compact, round, and separated by considerable gaps. There are no noise points, and all the names have been easily assigned to clusters.

```{r fig-1-clusters, fig.cap="**Fig 1.** Law firm names clusters resulting from UMAP → DBSCAN processing of law firm names embeddings. Numbers are assigned automatically by DBSCAN", fig.height=4}
clusters_plot
```

```{r table-1-clusters}
kable(
  clusters_info_table,
  col.names = c(
    "# of cluster", "# of names", "Description",
    "Category", "Strategy"
  )
)
```

Table 1 shows cluster sizes and sample names from the clusters with manually assigned short cluster descriptions, categories and strategies. The quantity of clusters (52) is relatively big, and their size is not even: while the largest group of names consists of items, the smallest one includes just 12 objects. Almost 1/3 of clusters (17) is comprised of a mixture of short (usually a single-word) names that cannot be given a reasonable description. They are marked as “Misc” and classified as “No-strategy”.

All the other names are split between three primary strategies. “Market” strategy is comprised of firms with names having generally positive connotations in Russian language that help to pursue a potential client in a high quality of legal aid and thus act as a some sort of advertisement (“Guarantor”, “Expert”, “Partnership”, “Helper”, "Center” categories and some others). “Belonging” strategy, as its name suggests, make a strong emphasis on their belonging to the legal services market by using the words “law”, “lawyer”, “legal” or various Russian equivalents of “law firm” phrase. A distinct category of law firms with this strategy is formed by organisations that use a personal branding that follow the broad pattern “(law firm) <last name of a person> with partners” where (law firm) is an optional part, and <last name of a person> is a pointer to a featured lawyer of this organisation. Even if “law firm” subpart of the firm name is optional, the entire pattern is strongly associated in Russia with lawyers, thus this type of firm naming marks the belonging to the legal services sphere. Finally, “Uniqueness” strategy encompasses companies with names mentioning a particular kind of legal services (e.g. “Consulting”), or a target customers group (e.g. “Business”), or a related area of specialization (real estate, finance, audit, development), or a specific geographic location.

```{r table-2-strategies-summary}
kable(
  strategy_counts,
  digits = c(NA, 0, 0),
  col.names = c("Strategy", "# of names", "%")
)
```

Among the strategies, “No-strategy” (i.e. the absence of a clear strategy) is the most popular (see Table 2). It encompasses more than 1/3 of all the studied names. Belonging strategy is the second by the popularity holding 30% of law firms. Uniqueness strategy is employed by approximately 20% of companies, and the least typical one, market strategy, is used by slightly more than 10% of law firms.

## 3.2. Spatial regularities in law firms naming

As Tables 3 and 4 demonstrate, there are few law firm names clusters with some regularities in the spatial distribution of law firm names. These two tables display results only for the clusters where global Moran's I coefficient is statistically significant. The values of Moran's I metric for all the other clusters are close to zero, and their adjusted p-value is above the selected threshold of 0.05, so the results for these clusters are not shown for brevity.

```{r table-3-moran-i-by-neighborhood}
moran_test_by_regions %>% 
  filter(p_adj < 0.05) %>% 
  kable(
    digits = c(0, NA, 2, 3),
    col.names = c("# of cluster", "Description", "Moran's I", "Adj. p-value")
  )
```

```{r table-4-moran-i-by-economic-regions}
moran_test_by_er %>% 
  filter(p_adj < 0.05) %>% 
  kable(
    digits = c(0, NA, 2, 3),
    col.names = c("# of cluster", "Description", "Moran's I", "Adj. p-value")
  )
```

According to the autocorrelation analysis results in both branches, there are in total four law firm names clusters where the spatial distribution is different from the random. One of these clusters, “Center or company with geographical location”, is present in both tables and is characterized by the highest value of Moran's I, that is about 0.66. Thus, it is likely that law firms with names mentioning a geographical location of the country are not distributed randomly within the country, but situated in a number of neighbouring regions. Two maps on the Fig. 2, despite several differences, demonstrate that it is the regions in Eastern and Western Siberia that are characterized by the highest concentration of “geographically referenced” law firm names. A brief manual investigation uncovers that such names usually reference a particular region (e.g. Altai), or a specific city (e.g. Krasnoyarsk), or entire geographical area as a whole (Siberia). This argues for the presence of some sort of local or regional identity which is of highest expression in Siberia. It is possible that similar geo-based naming is present in other parts of the country, but the automatic clusterisation of law firm names using their embeddings cannot handle them properly — either because the quantity of geolocated names in other parts of the country is substantially smaller or because the text vectors do not reflect the similarity between the Siberian geographical mentions and geographical mentions in the other parts of Russia. Overall, an issue of geolocated law firm names is intriguing and requires further investigation.

As for other clusters with statistically significant global Moran's I (“Forensic” names, “Experts and advisors”, and personal-branding pattern), the hotspots are situated in the European South of Russia. In the case of “Forensic” names, it is likely to be attributed to a tiny set of national republics of North Caucasus that are probably outliers, because their overall number of law firms and the distribution of naming strategies is substantially different from the one in the majority of regions. “Experts and advisors” cluster demonstrates hotspots in Rostov, Volgograd and Astrakhan regions. They are characterized by a relatively high number of law firms and thus cannot be left to as just outliers. However, the exact reasons for the spatial concentration of this naming pattern in these regions remains unclear. Finally, a personal branding cluster represents a drastic split of the European South into two groups of regions: the first one comprising of several national republics with high level of personal branding names and the second one consisting of some other national republics and all the regular regions with small share of such names. This may be seen as a slight argument for a position that European South of Russia is a conglomerate of two quite distinct groups of regions.

# 4. Conclusion

The findings of the study are somewhat contradictory. On the one hand, several hypotheses have been proven. These include the following: law firm names in neighboring regions are more similar than law firm names in non-neighboring regions; the smaller the geographic distance between regions, the more similar the law firm names; and law firm names within an economic region are typically more similar compared to the law firm names outside an economic region. On the other hand, the difference in the metrics of similarity is relatively small in each case, as is the correlation coefficient. It is therefore important to note that although the postulated phenomena do indeed exist, their magnitude is weak.

The three principal strategies for naming law firms are as follows: “Form”, “Law”, and “Service”. The first of these represents a minority approach, while the remaining majority is divided between two other strategies. In the European South of Russia and in the Volga economic region, “Service” is a dominant strategy. However, no other spatial patterns in law firms naming have been identified. Economic regions are typically characterized by a mixture of naming strategies, except for Volga-Vyatka, where the “Law” strategy prevails, and Volga and North Caucasus, where the “Service” strategy is dominant.

The contribution of this study is primarily theoretical and methodological in nature. It underscores the significance of firm names as a source of information for business- and economy-related research. It also proposes a technique for comprehensive analysis of branding in a geographical context through the use of modern natural language processing (NLP) solutions based on artificial neural networks. Additionally, it advocates for the utilization of disaggregated administrative open data for a spatial analysis. The findings demonstrate the effect of geographical proximity, neighboring, and economic relations on the similarity of the economic and, to a certain extent, cultural properties of the territories represented in firm names. The results generally align with the widely accepted classification of economic regions in Russia, although they highlight a few instances where a theoretically undesired internal heterogeneity within an economic region is observed. Finally, the study provides an illustration of a semi-automated brand classification within a specific economic sector and discovers a geographical variation in naming strategies that could be further investigated.

# Data availability

A dataset about Russian law firms, vectors of law firm names obtained from Yandex Foundation models, results of manual labelling, additional files used for mapping, as well as all the code used to conduct the study and generate the paper file are available at Zenodo: https://doi.org/10.5281/zenodo.13691418.

# References

<div id="refs"></div>