### 2025-01-28 law-firms-names/Paper.Rmd
Edit introduction and some parts of methods, #4

---
title: "Identification and Spatial Analysis of Law Firms Naming Patterns in Russia Using Text Embeddings"
author: "Pavel Syomin"
output:
  word_document:
    reference_doc: "journal-article/reference.docx"
bibliography: "journal-article/refs.bib"
csl: "journal-article/apa.csl"
link-bibliography: False  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 4.96, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples
```

```{r package-imports, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(here)
library(knitr)
library(readr)
library(readxl)
library(sf)
library(stringr)
library(tidyr)
library(units)
```

```{r data-imports, cache=TRUE, include=FALSE}
source(here("journal-article", "Script.R"))
```

<div custom-style="Affiliation">

Perm State University, Geographical faculty, Department of social and economic geography, Perm, Russia

E-mail: ntsp\@ya.ru

ORCID: 0000-0002-4015-9206

</div>

<div custom-style="Abstract">

**Abstract:** The present study analyzes law firm naming in Russia, with a particular focus on the geographical aspects of this phenomenon. The primary objective of this research is to ascertain the impact of the spatial relations between the country's regions, specifically those of a neighboring, proximity, and belonging to a parent-level subdivision, on the naming of law firms. To conduct the analysis, data regarding law firms is extracted from the open dumps of the state registry of small and medium-sized businesses. The names of the firms are transformed into numerical vectors, subsequently clustered with the KMeans method, and then manually grouped to identify naming strategies. Numerical representations of law firm names within each region are averaged to get region embeddings, which are then used to compare the names of law firms in different regions. The research findings indicate that regions situated closer to one another or sharing a common border tend to adopt slightly more similar law firm names. Furthermore, regions that are part of the same economic region also display a greater degree of similarity in law firm naming. Three primary strategies of law firm naming are “Form”, “Law“, and “Service”. Volga and North Caucasus economic regions are characterized by the predominance of the “Service” strategy, whereas the Volga-Vyatka economic region exhibits a “Law” strategy, and the remaining territories demonstrate a combination of naming strategies. The results of the study lend support to the widely used classification of economic regions in Russia. The findings also underscore the impact of neighboring, proximity, and interconnectivity on the economic and cultural attributes of the regions. The administrative open data processing techniques, as well as the methods of working with firm names based on neural networks proposed in this paper, may be used by other economic geographers for their own research purposes.

</div>

<div custom-style="Keywords">

**Keywords:** firm names; law firms; text embeddings; spatial analysis, autocorrelation, naming strategies.

</div>

# Statements and Declarations

**Competing Interests.** No funding was received for conducting this study. The author has no financial or proprietary interests in any material discussed in this article.

\newpage

# 1. Introduction

The geography of services, also known as the geography of the service sector, is a crucial component of geographical studies in Russia. This field has a history spanning several decades, with origins dating back to the 1960s. The geography of services has two distinct dimensions: economic and social. The former is primarily concerned with the business aspects of the services, while the latter addresses the implications for human well-being arising from the geographical organization of services [@Tkachenko2016]. The existing scientific literature on the geography of services in Russia covers a broad range of topics, including education [@Shulgina2019], car rental [@Zyrianova2018], catering [@Preobrazhenskiy2018], medical care [@Kazantcev2022; @Panin2023; @Nikoloski2024], and internet access [@Nagirnaya2015; @Zemtsov2022]. However, some areas remain underexplored, with legal services being one such example [@Tkachenko2016; @Moiseeva2016].

Despite the lack of interest from geographers, the spatial aspects of the activity of law firms are worth investigating for a number of reasons. Firstly, legal services can be categorized as business or producer services, and this category of services is important for economic development, on both a national and regional scale [@Hansen1993; @Faulconbridge2019]. Secondly, despite their comparatively minor direct economic impact, these services play a significant role as intermediaries and facilitators, contributing to the establishment and maintenance of the rule of law, social order, stability, and predictability of social relations. Thirdly, for the general public, access to legal aid is vital for the protection of human rights and the attainment of social justice. In this regard, the spatial aspects of law services are on par with the spatial issues of medical care [@Blacksell1988; @Blacksell1990].

The study of firm names — which are sometimes also referred to as company names, brand names, or simply “brands” — is a relatively uncommon yet potentially fruitful approach to the geographical analysis of services and related problems. A firm name is designed to distinguish the business from competitors, to highlight the sector or geographical area in which it operates, and to convey information about the services offered by a provider and the associated benefits. A brand is therefore capable of transferring a substantial amount of information in a brief textual format. Such information may prove useful in uncovering the city's business landscape and consumer perceptions of a particular city with its distinctive services [@Raento2001]. It may also help to analyze trends common to a specific sector in a geographic region [@Marjavaara2022]. The incorporation of place names (toponyms) in brands influences the business development and consumer attitudes [@Peluso2012; @Zhang2014; @Erdogan2016; @Noorda2019]. Company names encapsulate a multitude of intertwined properties relevant to the economy and culture of a given location. This synthesis makes them suitable for the identification of regions, especially in the case of vernacular regions [@Reed1976; @Liesch2015; @Vukosav2016; @Holtkamp2018].

Law firm names can follow different patterns. A naming pattern can be defined as as a typical way of branding, identified by the use of identical or semantically similar words or phrases. For instance, the names “Law firm “Femida” and “Law company “Alibi” follow one pattern, while “Centre of legal aid and support” is an example of another. Patterns can be grouped several times to form a hierarchical classification [@Anora2015]. Such a classification can help to identify naming strategies. Compared to patterns, strategies are broader groups of names, usually based on a common purpose, e.g. highlighting a company's affiliation to a certain activity (retail, banking, transport, education, etc). @BassolsiGardella2023 propose a two-dimensional Strategic Naming Model (SNM) that includes two pairs of opposing strategies: belonging and uniqueness, no-strategy and market. They discover a difference between three Colombian cities in the naming strategies of their tourism companies. It is likely that such a variation is not be limited to the cities of Colombia and the tourism industry and can also be found in the case of law firms in Russia.

The aim of this research is twofold. First, it attempts to identify the naming patterns of law firms in Russia and to categorise them according to the SNM framework. Second, it seeks to test whether there are any regularities in the spatial distribution of these patterns. The potential viability of these two propositions cannot be derived directly from the existing research, but is based on common sense observations of the similarities and differences of law firm naming in and between different regions of the country. It is highly unlikely that the names of law firms in Russia are so diverse as to be impossible to classify. Their geographical distribution is a much more difficult issue. At first glance, there are no obvious factors that could lead lawyers in a particular region to choose a common naming pattern for their firms. However, according to the famous Tobler's First Law of Geography, “near things are more related than distant things” [@Tobler1970], there is a possibility that law firm naming patterns may form spatial clusters and hotspots that reflect, at least to some extent, economic and cultural similarities of the territories in which these firms are located. There is an evidence of regional variation of branding in food industry [@Goroshko2022] and wine production [@Klimenko2023] in Russia. Even though the market of legal services seems to be far less dependant on the properties of place, the question of spatial regularities in naming patterns is also intriguing and worth testing.

# 2. Materials and methods

## 2.1. Background

## 2.2. Data

The issue of obtaining information about firm names is a challenging one. All Russian organizations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data resell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized enterprises (SME registry) may be used [@SMERegistry2024]. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. As a rule of thumb, all companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SME registry, without any action being required on their part. There are additional qualifications for the SME registry, such as the absence of non-SME or foreign control, but these are unlikely to affect a significant proportion of law firms. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SMB registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organisations and individuals.

The downloadable version of the registry includes, in particular, firm names, addresses of registration (up to a settlement), and codes of activity kinds compatible with NACE Rev. 2 (European Classification of Economic Activities). It should be noted that the registry does not include several types of firms such as foreign commercial law firms without a juridical person registered under Russian law, non-profit organisations, bar (advocacy) formations, large enterprises. As a result, some data are missing. However, for a number of reasons, this problem should be considered insignificant. Foreign law firms usually register their affiliated juridical person in Russia and are therefore not omitted from the SME registry. The share of non-profit organisations in the Russian legal services market does not exceed 2% [@Moiseeva2016]. The names of bar (advocacy) formations are usually quite uniform. They contain the kind of the formation (advocate's office, advocate's bureau, college of advocates, juridical consultancy) accompanied by the personal name(s) of advocates or a reference to a geographical location. Their naming patterns are thus quite apparent and do not require further investigation. Large enterprises represent only a tiny minority of law firms.

For a validation, a list of law firms' unique tax IDs obtained from the SME registry for the year 2023 is matched with the list of law firms' tax IDs from the Russian Financial Statements Database (RFSD) [@Bondarkov2025] for the same year. The comparison shows that ~3,300 of the firms present in the SME registry are missing from the RFSD, and ~5,200 firms present in the RFSD are missing from the SME registry. The balance of missing records is not so far from being equal. The SME registry contains ~34000 law firms for the year 2023, so the missing rate is below 15%. Consequently, the completeness of the data is relatively high and reliable conclusions can be drawn.

## 2.2. Methods

### 2.2.1. Law firms selection criteria

In this study, the term “law firm” is defined as a commercial organization registered in Russia in accordance with the Civil Code of Russia, with a NACE Rev. 2-compatible main activity code equal to 69.10 (activity in the area of law). The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, bar (advocacy) formations, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SME registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

The data used in the study cover a period of six years, from 2016 to 2021. The initial point in time is defined by the commencement of SME registry data availability. The end point, 2021, is the last year of “normal” conditions in Russia unaffected by the military operation in Ukraine and the resulting international sanctions and domestic economic and political turmoil. The purpose of the time dimension is to rule out short-lived law firms, which are likely to be shell companies. Organisations with a lifespan of less than three years are excluded from the research, while all the remaining are included in the further analysis. In addition, the use of multi-year data acts as a form of averaging, smoothing out the potential biases in the spatial structure caused by random annual changes in the population of active law firms.

### 2.2.2. Firm data acquisition

The processing of SME registry data is conducted in a series of stages. In the initial phase of the research process, all source archives are downloaded. Each source archive contains a full copy of the registry data for the specified publication date, with a new archive becoming available on a monthly basis. Subsequently, data pertaining to law firms is extracted from the archives and stored in CSV files. Each CSV file contains data from one source archive. Subsequently, all CSV files are consolidated into a single table, the duplicated records are removed, and the remaining records are augmented with start and end temporal markers. Subsequently, the addresses are normalized to resolve the inconsistencies in location names and geocoded up to a city/settlement level to add geographical coordinates. The lack of within-city location in the source data prevents from more fine-grained geocoding but does not pose a serious problem because city-level coordinates are enough for large-scale analysis on country or region level. Finally, the resulting table is transformed into a panel view, with each row representing a single law firm at a given year. All these stages are implemented in an open-source Python command-line tool.

### 2.2.3. Firm name embeddings retrieval

The Yandex Foundation Models cloud service [@Yandex2024] is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organizational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed. The dimension of a firm name vector is 256. Contextual embeddings (YandexGPT) were used in preference to static ones (e.g. FastText), because in the preliminary analysis, they have resulted in a much better clustering of law firm names, which is crucial for naming patterns identification.

### 2.2.4. Naming patterns identification and classification

Law firm names are clustered using the KMeans technique. A matrix of features for clustering is constructed from the name embeddings. As there are 256 elements in each vector, the feature matrix has 256 columns. A target number of clusters is determined with the elbow method and ultimately set to 50, with the understanding that further manual processing of the clustering results will be necessary. Each name is assigned a cluster label between 1 and 50. Despite the simplicity of the clustering algorithm, which can even be considered old-fashioned compared to (H)DBSCAN, it performs quite well on the data at hand. Preliminary tests have shown that the clusters identified by (H)DBSCAN are not significantly better than those obtained by KMeans, but the latter takes much less times and memory. It is important to emphasise that automatic clustering is only a auxiliary step in the identification and classification of naming strategies which aims to facilitate the manual data processing by breaking down the entire set of law firm names into the compact groups that can be analysed more easily by a human. Therefore, there is no reason to improve the quality of clustering too much.

The resulting clusters are then manually grouped. A random sample of 20 names is selected from each cluster. With 50 clusters and 20 sampled names allocated in each, the total size of the dataset for manual processing is 1, 000. The cluster samples are carefully examined, and a concise description is provided for each, e.g. “Names with the word “консалт/консалтинг” (consult/consulting)”. After this, the clusters are examined again and assigned to a naming pattern, e.g. “Consulting” or “Law firm”. Finally, the clusters and their naming patterns are divided into four previously mentioned naming strategies derived from @BassolsiGardella2023. Each cluster belongs to a single pattern, and each pattern belongs to a single strategy. Each strategy consists of several patterns, and each pattern consists of one or many clusters, forming a hierarchical structure.

Each group except for the “Misc” category represents a specific naming strategy. The “Form” strategy encompasses names that are based on a word or words denoting a kind of business entity, including “company”, “firm”, “group”, “partnership”, “center”, “alliance”, and “corporation”. The “Law” strategy emphasizes the firm's affiliation with the legal services market, employing terms such as “legal”, “juridical”, “lex”, “law”, and “lawyer”. Names belonging to the “Service” strategy may indicate a particular service provided by a law firm (“advisor”, “consulting”, “protection”, “support”, “warrant”), describe the sectors of businesses the firm is primarily engaged with (“technology”, “development”, “house management”, “accounting”), or mention services in general. The “Misc” category is a catch-all group for names that do not clearly align with any other category.

### 2.2.5. Spatial regularities identification

Analysis of autocorrelation is used to check whether the spatial distribution of names belonging to each cluster identified on the previous step is random or has some regularities. Regions (constituent entries) or the Russian Federation were selected as a spatial subdivision for calculating the autocorrelation measures. It could have been more reasonable to use the next spatial level, municipal territories, to perform the computations, but exploratory data analysis has shown that a lion's share of such territories has no law firms, because the legal business is highly concentrated in cities and large settlements.

Six spatial weights matrices are used to ensure the robustness of the results. Three of these matrices are constructed with K nearest neighbours technique based on the location of region's centroids, where K equals 3, 4 or 6, respectively. Another matrix also uses geographical properties of regions, but instead of centroids, it relies on the existence of common border lines or points to identify neighbouring regions. Sakhalin and Kaliningrad regions, being enclaves and thus having zero immediate neighbors, were manually connected with a few closest regions. Two remaining matrices treat regions belonging to the same upper-level metaregion as neighbours, and all the other regions as non-neighbours. For one of these two matrices, economic regions act as metaregions, and for another, federal districts are used in this fashion. Economic regions are defined according to the classification widely used by economic geographers and spatial development practitioners in Russia for decades and outlined, for example, in [@Leyzerovich2010]. Federal districts composition is defined in the Executive Order of the President of Russia No. 849, Dated May 13, 2000, “On the Plenipotentiary Representative of the President of the Russian Federation in the Federal District” (with up-to-date amendments).

For each cluster, a global Moran's I metric is calculated. The variable of interest is a share of law firm names belonging to the cluster in a total quantity of law firm names in a region. If there are no law firm names of the cluster in the region, the share is considered to equal zero. P-values are adjusted for multiple comparisons with Holm's correction. Significance level (alpha) is set to 0.05.

For the clusters with statistically significant global Moran's I, a local Moran's I was subsequently calculated. P-values are adjusted with local Holm's correction. Cluster classification of local Moran's I are mapped.

### 2.2.6. Data statistics

There are `r initial_firms_count` law firms in the SME registry in the selected period with `r nrow(unique_names)` unique names. From the initial dataset, `r valid_firms_count` law firms have existed for at least three years and have been selected for the analysis. Their names are sometimes duplicated, and the number of unique law firm names being analysed is `r length(unique(clustered$name))`.

# 3. Results and discussion

## 3.1. Law firms naming strategies

Law firm names clusters resulted from UMAP → DBSCAN pipeline are illustrated using scatterplot in Fig. 1. The visual quality of clusterisation is high: all the clusters are compact, round, and separated by considerable gaps. There are no noise points, and all the names have been easily assigned to clusters.

```{r fig-1-clusters, fig.cap="**Fig 1.** Law firm names clusters resulting from UMAP → DBSCAN processing of law firm names embeddings. Numbers are assigned automatically by DBSCAN", fig.height=4}
clusters_plot
```

```{r table-1-clusters}
kable(
  clusters_info_table,
  col.names = c(
    "# of cluster", "# of names", "Description",
    "Category", "Strategy"
  )
)
```

Table 1 shows cluster sizes and sample names from the clusters with manually assigned short cluster descriptions, categories and strategies. The quantity of clusters (52) is relatively big, and their size is not even: while the largest group of names consists of items, the smallest one includes just 12 objects. Almost 1/3 of clusters (17) is comprised of a mixture of short (usually a single-word) names that cannot be given a reasonable description. They are marked as “Misc” and classified as “No-strategy”.

All the other names are split between three primary strategies. “Market” strategy is comprised of firms with names having generally positive connotations in Russian language that help to pursue a potential client in a high quality of legal aid and thus act as a some sort of advertisement (“Guarantor”, “Expert”, “Partnership”, “Helper”, "Center” categories and some others). “Belonging” strategy, as its name suggests, make a strong emphasis on their belonging to the legal services market by using the words “law”, “lawyer”, “legal” or various Russian equivalents of “law firm” phrase. A distinct category of law firms with this strategy is formed by organisations that use a personal branding that follow the broad pattern “(law firm) <last name of a person> with partners” where (law firm) is an optional part, and <last name of a person> is a pointer to a featured lawyer of this organisation. Even if “law firm” subpart of the firm name is optional, the entire pattern is strongly associated in Russia with lawyers, thus this type of firm naming marks the belonging to the legal services sphere. Finally, “Uniqueness” strategy encompasses companies with names mentioning a particular kind of legal services (e.g. “Consulting”), or a target customers group (e.g. “Business”), or a related area of specialization (real estate, finance, audit, development), or a specific geographic location.

```{r table-2-strategies-summary}
kable(
  strategy_counts,
  digits = c(NA, 0, 0),
  col.names = c("Strategy", "# of names", "%")
)
```

Among the strategies, “No-strategy” (i.e. the absence of a clear strategy) is the most popular (see Table 2). It encompasses more than 1/3 of all the studied names. Belonging strategy is the second by the popularity holding 30% of law firms. Uniqueness strategy is employed by approximately 20% of companies, and the least typical one, market strategy, is used by slightly more than 10% of law firms.

## 3.2. Spatial regularities in law firms naming

As Tables 3 and 4 demonstrate, there are few law firm names clusters with some regularities in the spatial distribution of law firm names. These two tables display results only for the clusters where global Moran's I coefficient is statistically significant. The values of Moran's I metric for all the other clusters are close to zero, and their adjusted p-value is above the selected threshold of 0.05, so the results for these clusters are not shown for brevity.

```{r table-3-moran-i-by-neighborhood}
moran_test_by_regions %>% 
  filter(p_adj < 0.05) %>% 
  kable(
    digits = c(0, NA, 2, 3),
    col.names = c("# of cluster", "Description", "Moran's I", "Adj. p-value")
  )
```

```{r table-4-moran-i-by-economic-regions}
moran_test_by_er %>% 
  filter(p_adj < 0.05) %>% 
  kable(
    digits = c(0, NA, 2, 3),
    col.names = c("# of cluster", "Description", "Moran's I", "Adj. p-value")
  )
```

According to the autocorrelation analysis results in both branches, there are in total four law firm names clusters where the spatial distribution is different from the random. One of these clusters, “Center or company with geographical location”, is present in both tables and is characterized by the highest value of Moran's I, that is about 0.66. Thus, it is likely that law firms with names mentioning a geographical location of the country are not distributed randomly within the country, but situated in a number of neighbouring regions. Two maps on the Fig. 2, despite several differences, demonstrate that it is the regions in Eastern and Western Siberia that are characterized by the highest concentration of “geographically referenced” law firm names. A brief manual investigation uncovers that such names usually reference a particular region (e.g. Altai), or a specific city (e.g. Krasnoyarsk), or entire geographical area as a whole (Siberia). This argues for the presence of some sort of local or regional identity which is of highest expression in Siberia. It is possible that similar geo-based naming is present in other parts of the country, but the automatic clusterisation of law firm names using their embeddings cannot handle them properly — either because the quantity of geolocated names in other parts of the country is substantially smaller or because the text vectors do not reflect the similarity between the Siberian geographical mentions and geographical mentions in the other parts of Russia. Overall, an issue of geolocated law firm names is intriguing and requires further investigation.

As for other clusters with statistically significant global Moran's I (“Forensic” names, “Experts and advisors”, and personal-branding pattern), the hotspots are situated in the European South of Russia. In the case of “Forensic” names, it is likely to be attributed to a tiny set of national republics of North Caucasus that are probably outliers, because their overall number of law firms and the distribution of naming strategies is substantially different from the one in the majority of regions. “Experts and advisors” cluster demonstrates hotspots in Rostov, Volgograd and Astrakhan regions. They are characterized by a relatively high number of law firms and thus cannot be left to as just outliers. However, the exact reasons for the spatial concentration of this naming pattern in these regions remains unclear. Finally, a personal branding cluster represents a drastic split of the European South into two groups of regions: the first one comprising of several national republics with high level of personal branding names and the second one consisting of some other national republics and all the regular regions with small share of such names. This may be seen as a slight argument for a position that European South of Russia is a conglomerate of two quite distinct groups of regions.

# 4. Conclusion

The findings of the study are somewhat contradictory. On the one hand, several hypotheses have been proven. These include the following: law firm names in neighboring regions are more similar than law firm names in non-neighboring regions; the smaller the geographic distance between regions, the more similar the law firm names; and law firm names within an economic region are typically more similar compared to the law firm names outside an economic region. On the other hand, the difference in the metrics of similarity is relatively small in each case, as is the correlation coefficient. It is therefore important to note that although the postulated phenomena do indeed exist, their magnitude is weak.

The three principal strategies for naming law firms are as follows: “Form”, “Law”, and “Service”. The first of these represents a minority approach, while the remaining majority is divided between two other strategies. In the European South of Russia and in the Volga economic region, “Service” is a dominant strategy. However, no other spatial patterns in law firms naming have been identified. Economic regions are typically characterized by a mixture of naming strategies, except for Volga-Vyatka, where the “Law” strategy prevails, and Volga and North Caucasus, where the “Service” strategy is dominant.

The contribution of this study is primarily theoretical and methodological in nature. It underscores the significance of firm names as a source of information for business- and economy-related research. It also proposes a technique for comprehensive analysis of branding in a geographical context through the use of modern natural language processing (NLP) solutions based on artificial neural networks. Additionally, it advocates for the utilization of disaggregated administrative open data for a spatial analysis. The findings demonstrate the effect of geographical proximity, neighboring, and economic relations on the similarity of the economic and, to a certain extent, cultural properties of the territories represented in firm names. The results generally align with the widely accepted classification of economic regions in Russia, although they highlight a few instances where a theoretically undesired internal heterogeneity within an economic region is observed. Finally, the study provides an illustration of a semi-automated brand classification within a specific economic sector and discovers a geographical variation in naming strategies that could be further investigated.

# Data availability

A dataset about Russian law firms, vectors of law firm names obtained from Yandex Foundation models, results of manual labelling, additional files used for mapping, as well as all the code used to conduct the study and generate the paper file are available at Zenodo: https://doi.org/10.5281/zenodo.13691418.

# References

<div id="refs"></div>