### 2025-01-28 law-firms-names/Paper.Rmd
Edit methods, results, discussion, conclusion, figures, abstract, and style, before submitting the second revision, #4

---
title: "Identification and Spatial Analysis of Law Firms Naming Patterns in Russia Using Text Embeddings"
author: "Pavel Syomin"
output:
  word_document:
    reference_doc: "journal-article/reference.docx"
bibliography: "journal-article/refs.bib"
csl: "journal-article/apa_numeric.csl"
link-bibliography: False  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 6.7, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples
```

```{r package-imports, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(here)
library(knitr)
library(readr)
library(readxl)
library(sf)
library(stringr)
library(tidyr)
library(units)
```

```{r data-imports, cache=TRUE, include=FALSE}
source(here("journal-article/Script.R"))
```

<div custom-style="Affiliation">

Perm State University, Geographical faculty, Department of social and economic geography, Perm, Russia

E-mail: ntsp\@ya.ru

ORCID: 0000-0002-4015-9206

</div>

<div custom-style="Abstract">

**Abstract:** The present study analyzes the naming of law firms in Russia, with a particular focus on the geographical aspects of this phenomenon. The main objectives of this research are to identify law firm naming patterns, to categorise them and to test whether there are any spatial regularities in the distribution of law firms according to the identified patterns. To conduct the analysis, data regarding law firms is extracted from the open dumps of the state registry of small and medium-sized enterprises. The names of the firms are transformed into numerical vectors, subsequently clustered with the KMeans method, and then manually grouped to identify naming patterns and strategies. Global and local Moran's *I* statistics are calculated to test for spatial autocorrelation in the proportion of a given pattern in the regions. As a result, 22 naming patterns are identified, grouped into four naming strategies. Apart from no-strategy (40% of law firms), the most popular strategy is belonging (to the legal services market — 30%), followed by uniqueness (of service or location — 20%) and market (words and phrases with positive or promotional meaning — 10%). The research findings indicate that the spatial distribution of law firm naming patterns is generally not significantly different from random, with two exceptions: geographical branding in Siberia and an unusual heighborhood of regions with high and low proportions of “Expert” and “Personal brand” patterns in the European South.

</div>

<div custom-style="Keywords">

**Keywords:** firm names; law firms; text embeddings; spatial analysis, autocorrelation, naming strategies.

</div>

# Statements and Declarations

**Competing Interests.** No funding was received for conducting this study. The author has no financial or proprietary interests in any material discussed in this article.

\newpage

# 1. Introduction

The geography of services, also known as the geography of the service sector, is a crucial component of geographical studies in Russia. This field has a history spanning several decades, with origins dating back to the 1960s. The geography of services has two distinct dimensions: economic and social. The former is primarily concerned with the business aspects of the services, while the latter addresses the implications for human well-being arising from the geographical organization of services [@Tkachenko2016]. The existing scientific literature on the geography of services in Russia covers a broad range of topics, including education [@Shulgina2019], car rental [@Zyrianova2018], catering [@Preobrazhenskiy2018], medical care [@Kazantcev2022; @Panin2023; @Nikoloski2024], and internet access [@Nagirnaya2015; @Zemtsov2022]. However, some areas remain underexplored, with legal services being one such example [@Tkachenko2016; @Moiseeva2016].

Despite the lack of interest from geographers, the spatial aspects of the activity of law firms are worth investigating for a number of reasons. Firstly, legal services can be categorized as business or producer services, and this category of services is important for economic development, on both a national and regional scale [@Hansen1993; @Faulconbridge2019]. Secondly, despite their comparatively minor direct economic impact, these services play a significant role as intermediaries and facilitators, contributing to the establishment and maintenance of the rule of law, social order, stability, and predictability of social relations. Thirdly, for the general public, access to legal aid is vital for the protection of human rights and the attainment of social justice. In this regard, the spatial aspects of law services are on par with the spatial issues of medical care [@Blacksell1988; @Blacksell1990].

The study of firm names — which are sometimes also referred to as company names, brand names, or simply “brands” — is a relatively uncommon yet potentially fruitful approach to the geographical analysis of services and related problems. A firm name is designed to distinguish the business from competitors, to highlight the sector or geographical area in which it operates, and to convey information about the services offered by a provider and the associated benefits. A brand is therefore capable of transferring a substantial amount of information in a brief textual format. Such information may prove useful in uncovering the city's business landscape and consumer perceptions of a particular city with its distinctive services [@Raento2001]. It may also help to analyze trends common to a specific sector in a geographic region [@Marjavaara2022]. The incorporation of place names (toponyms) in brands influences the business development and consumer attitudes [@Peluso2012; @Zhang2014; @Erdogan2016; @Noorda2019]. Company names encapsulate a multitude of intertwined properties relevant to the economy and culture of a given location. This synthesis makes them suitable for the identification of regions, especially in the case of vernacular regions [@Reed1976; @Liesch2015; @Vukosav2016; @Holtkamp2018].

Law firm names can follow different patterns. A naming pattern can be defined as as a typical way of branding, identified by the use of identical or semantically similar words or phrases. For instance, the names “Law firm “Femida” and “Law company “Alibi” follow one pattern, while “Centre of legal aid and support” is an example of another. Patterns can be grouped several times to form a hierarchical classification [@Arora2015]. Such a classification can help to identify naming strategies. Compared to patterns, strategies are broader groups of names, usually based on a common purpose, e.g. highlighting a company's affiliation to a certain activity (retail, banking, transport, education, etc). Bassols i Gardella et al. [@BassolsiGardella2023] propose a two-dimensional Strategic Naming Model (SNM) that includes two pairs of opposing strategies: belonging and uniqueness, no-strategy and market. They discover a difference between three Colombian cities in the naming strategies of their tourism companies. It is likely that such a variation is not be limited to the cities of Colombia and the tourism industry and can also be found in the case of law firms in Russia.

The aim of this research is twofold. First, it attempts to identify the naming patterns of law firms in Russia and to categorise them according to the SNM framework. Second, it seeks to test whether there are any regularities in the spatial distribution of these patterns. The potential viability of these two propositions cannot be derived directly from the existing research, but is based on common sense observations of the similarities and differences of law firm naming in and between different regions of the country. It is highly unlikely that the names of law firms in Russia are so diverse as to be impossible to classify. Their geographical distribution is a much more difficult issue. At first glance, there are no obvious factors that could lead lawyers in a particular region to choose a common naming pattern for their firms. However, according to the famous Tobler's First Law of Geography, “near things are more related than distant things” [@Tobler1970], there is a possibility that law firm naming patterns may form spatial clusters and hotspots that reflect, at least to some extent, economic and cultural similarities of the territories in which these firms are located. There is an evidence of regional variation of branding in food industry [@Goroshko2022] and wine production [@Klimenko2023] in Russia. Even though the market of legal services seems to be far less dependant on the properties of place, the question of spatial regularities in naming patterns is also intriguing and worth testing.

# 2. Materials and methods

## 2.1. Background

Russia belongs to the Germanistic branch of a civil law legal system. Legal services (advice, negotiation, representation in court, etc) for organisations and individuals are usually provided by private law companies and individual lawyers. The market is divided between ordinary lawyers and advocates (members of the Bar), but there is no bar monopoly. There are ~90, 000 law firms in Russia, including ~40, 000 commercial law companies, ~30, 000 sole practitioners, ~20, 000 bar (advocate) offices, and ~10, 000 bar (advocate) organisations. There are no geographical restrictions on legal practitioners, so that a law firm registered in one region of the country can provide services in any other region. There are full-service and specialised law firms, as well as Russian law firms (the majority) and foreign law firms (the influential minority).

## 2.2. Data

The issue of obtaining information about firm names is a challenging one. All Russian organizations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data resell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized enterprises (SME registry) may be used [@SMERegistry2024]. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. As a rule of thumb, all companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SME registry, without any action being required on their part. There are additional qualifications for the SME registry, such as the absence of non-SME or foreign control, but these are unlikely to affect a significant proportion of law firms. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SMB registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organisations and individuals.

The downloadable version of the registry includes, in particular, firm names, addresses of registration (up to a settlement), and codes of activity kinds compatible with NACE Rev. 2 (European Classification of Economic Activities). It should be noted that the registry does not include several types of firms such as foreign commercial law firms without a juridical person registered under Russian law, non-profit organisations, bar (advocacy) formations, large enterprises. As a result, some data are missing. However, for a number of reasons, this problem should be considered insignificant. Foreign law firms usually register their affiliated juridical person in Russia and are therefore not omitted from the SME registry. The share of non-profit organisations in the Russian legal services market does not exceed 2% [@Moiseeva2016]. The names of bar (advocacy) formations are usually quite uniform. They contain the kind of the formation (advocate's office, advocate's bureau, college of advocates, juridical consultancy) accompanied by the personal name(s) of advocates or a reference to a geographical location. Their naming patterns are thus quite apparent and do not require further investigation. Large enterprises represent only a tiny minority of law firms.

For a validation, a list of law firms' unique tax IDs obtained from the SME registry for the year 2023 is matched with the list of law firms' tax IDs from the Russian Financial Statements Database (RFSD) [@Bondarkov2025] for the same year. The comparison shows that ~3,300 of the firms present in the SME registry are missing from the RFSD, and ~5,200 firms present in the RFSD are missing from the SME registry. The balance of missing records is not so far from being equal. The SME registry contains ~34000 law firms for the year 2023, so the missing rate is below 15%. Consequently, the completeness of the data is relatively high and reliable conclusions can be drawn.

## 2.3. Methods

### 2.3.1. Law firms selection criteria

In this study, the term “law firm” is defined as a commercial organization registered in Russia in accordance with the Civil Code of Russia, with a NACE Rev. 2-compatible main activity code equal to 69.10 (activity in the area of law). The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, bar (advocacy) formations, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SME registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

The data used in the study cover a period of six years, from 2016 to 2021. The initial point in time is defined by the commencement of SME registry data availability. The end point, 2021, is the last year of “normal” conditions in Russia unaffected by the military operation in Ukraine and the resulting international sanctions and domestic economic and political turmoil. The purpose of the time dimension is to rule out short-lived law firms, which are likely to be shell companies. Organisations with a lifespan of less than three years are excluded from the research, while all the remaining are included in the further analysis. In addition, the use of multi-year data acts as a form of averaging, smoothing out the potential biases in the spatial structure caused by random annual changes in the population of active law firms.

### 2.3.2. Firm data acquisition

The processing of SME registry data is conducted in a series of stages. In the initial phase of the research process, all source archives are downloaded. Each source archive contains a full copy of the registry data for the specified publication date, with a new archive becoming available on a monthly basis. Subsequently, data pertaining to law firms is extracted from the archives and stored in CSV files. Each CSV file contains data from one source archive. Subsequently, all CSV files are consolidated into a single table, the duplicated records are removed, and the remaining records are augmented with start and end temporal markers. Subsequently, the addresses are normalized to resolve the inconsistencies in location names and geocoded up to a city/settlement level to add geographical coordinates. The lack of within-city location in the source data prevents from more fine-grained geocoding but does not pose a serious problem because city-level coordinates are enough for large-scale analysis on country or region level. Finally, the resulting table is transformed into a panel view, with each row representing a single law firm at a given year. All these stages are implemented in an open-source Python command-line tool.

### 2.3.3. Firm name embeddings retrieval

The Yandex Foundation Models cloud service [@Yandex2024] is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organizational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed. The dimension of a firm name vector is 256. Contextual embeddings (YandexGPT) were used in preference to static ones (e.g. FastText), because in the preliminary analysis, they have resulted in a much better clustering of law firm names, which is crucial for naming patterns identification.

### 2.3.4. Naming patterns identification and classification

Law firm names are clustered using the KMeans technique. A matrix of features for clustering is constructed from the name embeddings. As there are 256 elements in each vector, the feature matrix has 256 columns. A target number of clusters is determined with the elbow method and ultimately set to 50, with the understanding that further manual processing of the clustering results will be necessary. Each name is assigned a cluster label between 1 and 50. Despite the simplicity of the clustering algorithm, which can even be considered old-fashioned compared to (H)DBSCAN, it performs quite well on the data at hand. Preliminary tests have shown that the clusters identified by (H)DBSCAN are not significantly better than those obtained by KMeans, but the latter takes much less times and memory. It is important to emphasise that automatic clustering is only a auxiliary step in the identification and classification of naming strategies which aims to facilitate the manual data processing by breaking down the entire set of law firm names into the compact groups that can be analysed more easily by a human. Therefore, there is no reason to improve the quality of clustering too much.

The resulting clusters are then manually grouped. A random sample of 20 names is selected from each cluster. With 50 clusters and 20 sampled names allocated in each, the total size of the dataset for manual processing is 1, 000. The cluster samples are carefully examined, and a concise description is provided for each, e.g. “Names with the word “консалт/консалтинг” (consult/consulting)”. After this, the clusters are examined again and assigned to a naming pattern, e.g. “Consulting” or “Law firm”. Finally, the clusters and their naming patterns are divided into four previously mentioned naming strategies derived from Bassols i Gardella et al. [@BassolsiGardella2023]. Each cluster belongs to a single pattern, and each pattern belongs to a single strategy. Each strategy consists of several patterns, and each pattern consists of one or many clusters, forming a hierarchical structure.

### 2.3.5. Testing for spatial regularities

Spatial autocorrelation analysis is used to check whether the spatial distribution of names following each pattern identified in the previous step is random or has some regularities. Regions (constituent entities) of the Russian Federation are chosen as the unit of observation. Municipal territories, the next level of spatial division, are not used, because an exploratory data analysis has shown that the lion's share of such the spatial units do not have any law firms.

Six spatial weight matrices are utilized to ensure the robustness of the results. Three of these matrices are constructed using the K-nearest neighbors technique, where K is 3, 4 or 6. Another matrix relies on the existence of common boundaries or points to detect neighboring regions. The Sakhalin and Kaliningrad regions, which are enclaves, are manually linked to a few nearest regions. Two remaining matrices treat regions belonging to the same upper-level macro-region as neighbors, and all other regions as non-neighbors. One of these matrices uses economic regions act as macro-regions, and the other uses federal districts. Economic regions are defined according to the classification that has been widely used by economic geographers and spatial development practitioners in Russia for decades, as outlined, for example, in [@Leyzerovich2010]. The composition of federal districts is defined by the Decree of the President of the Russian Federation (No. 849 of May 13, 2000) “On the Plenipotentiary Representative of the President of the Russian Federation in the Federal District”, with respective amendments.

A global Moran's *I* statistic [@Moran1948] is calculated for each combination of naming pattern and spatial weight matrix. The variable of interest is the proportion of law firm names that follow a pattern in a total set of law firm names in a region. If there are no law firm names of the pattern in the region, the proportion is considered to be zero. For those naming patterns where the global Moran's *I* is statistically significant, and this significance is robust, a local Moran's *I~i~* is calculated for each region. In addition, each region is categorised and labelled according to its value and the values of its neighbors. Four typical categories are used: “High-High” (high value in a region and its neighbors), “Low-Low” (an opposite), “Low-High”, “High-Low”. *P*-values are adjusted for multiple comparisons using Holm's correction [@Holm1979]. The significance level is set at 0.05.

### 2.3.6. Data statistics

There are `r initial_firms_count` law firms in the SME registry in the selected period with `r nrow(unique_names)` unique names. From the initial dataset, `r valid_firms_count` law firms have existed for at least three years and have been selected for the analysis. Their names are sometimes duplicated, and the number of unique law firm names being analysed is `r length(unique(clustered$name))`.

# 3. Results and discussion

## 3.1. Law firms naming patterns and strategies

After clustering and manual grouping, 22 naming patterns are identified and assigned to one of four naming strategies (see Fig. 1 for a dendrogram and Table 1 for a complete list of clusters, their descriptions, patterns, and strategies). The most popular is the no-strategy followed by ~18, 700 law firms. Slightly less important is the belonging strategy with ~15, 000 companies. The uniqueness strategy (~8, 300 law firms) is relatively rare, and the least used is the market strategy (~5, 300 companies).

Belonging strategy includes law firms that directly state that they are law firms. “Law firm” is the most common pattern of this strategy. Most of the others (“Law”, “Lawyer”, “Legal service”) are semantically similar to it. It may seem questionable to include the “Personal brand” pattern (names such as “Mr. Brilliant and partners”) in this category, since personal branding also serves as a differentiation tool and thus implies uniqueness with the corresponding strategy. However, in Russia this way of naming is associated with lawyers, and such names often include the mention of law (“Law firm of Mr. Brilliant” instead of just “Mr. Brilliant”). In addition, Bassols i Gardella et al. [@BassolsiGardella2023] also categorise the personal branding of tourism firms as the belonging strategy.

Market strategy, according to the original definition, is used by firms “having a name directly appealing to markets” [@BassolsiGardella2023]. In the case of law firms, this can be achieved by mentioning target clients (“Business” pattern) or by using words with positive meanings that attract customers (“Positive”, “Plus”, “Legal aid”, “Guarantor” patterns). The “Expert” pattern is complex, as this word can be used to indicate either a high level of service (“expert” as a synonym for “professional”) or forensic services (“expert” as a forensic scientist). It is included here into market strategy because the semantic similarity of “experts” and “forensic scientists” in the Russian language is probably used as a marketing tool by forensic companies.

Uniqueness strategy encompasses law firms whose names mention a particular type of service (“Property&Development”, “Consulting”, “Audit&Finance”) or a specific geographical region (“Siberia”). In contrast to belonging, these firms usually do not indicate in their names that they are part of legal services market. The opposition between belonging and uniqueness strategies may reflect a division between general and specialised law firms. Patterns of this strategy have the potential to be used to identify the specialisation of law firms, although this requires an additional research.

No-strategy is a catch-all group for names that do not fit into any of the other three. Its main component is miscellaneous one-word names that are difficult to categorise or describe. This strategy also includes names containing the words “Group”, “Partnership”, “Center” or “Service” (usually without any meaningful clarification of the activity or type of service). The “Managing companies” pattern is, after some consideration, classified as no-strategy. In Russia, the term “managing companies” may define either property or asset management firms or house management companies, and although some of their services may be treated as paralegal, it is difficult to place them in either the belonging, uniqueness, or market strategy The same logic applies to the “EdSciTech” pattern with probably education or LegalTech companies.

## 3.2. Spatial regularities in law firms naming

There are few patterns (3 out of 22) with statistically significant regularities in their spatial distribution (Fig. 2). These results are stable regardless of the technique used to construct the spatial weight matrix. The “Siberia” pattern is characterised by the highest global Moran's *I*, while the values for the “Expert” and “Personal brand” patterns are much lower, but still statistically significant.

A hotspot of the “Siberia” pattern is unsurprisingly located in Siberia (Fig. 3a). It includes four core regions (Novosibirsk, Tomsk, Kemerovo regions, Altai Territory) and a periphery of a few other regions (e.g. Omsk region, Republic of Khakassia, Republic of Altai). The hotspots of two other patterns (Fig. 3b, Fig. 3c) are located in the European South of Russia. This part of Russia is characterised by a mixture of regions where the popularity of a particular naming pattern is either high in a similar neighborhood or low in the neighborhood of regions with high values. This observation can be partly attributed to the presence of outliers among the national republics of the North Caucasus where the characteristics and composition of the legal services market differ significantly from their nearest neighbors. Some of these nearest neighbors (e.g. Krasnodar Territory) are, on the contrary, the regions with a relatively developed legal services market. In summary, the analysis and mapping of its results revealed two geographical parts of Russia where some naming patterns differ from the rest of the country: Siberia and the European South. However, the majority of naming patterns show no spatial regularities, i.e. their spatial distribution is not statistically significantly different from random.

# 4. Conclusion

The findings of the study are somewhat contradictory. On the one hand, there are 22 law firm naming patterns in Russia that can be attributed to one of the four broad SMN strategies: belonging or uniqueness, market or no-strategy. Apart from the no-strategy (~40% of firms), belonging is the most popular strategy used by 30% of law firms, followed by uniqueness (~20%) and market (~10%). On the other hand, in gerenal there are no spatial regularities in the distribution of law firms that use a particular pattern: all the parts of the country are generally the same when it comes to law firms naming. Two notable exceptions to this rule are Siberia, with a fairly popular geographical branding (“Siberia” pattern), and the European South, with an unusual spatial composition of “Expert” and “Personal brand” patterns.

The contribution of this study is primarily theoretical and methodological in nature. It underscores the significance of firm names as a source of information for business- and economy-related research. It also proposes a technique for comprehensive analysis of branding in a geographical context through the use of modern natural language processing (NLP) solutions based on artificial neural networks. Additionally, it advocates for the utilization of disaggregated administrative open data for a spatial analysis. The findings demonstrate that even though the legal services market looks quite disconnected from the geography, the place can still play some role in its branding. Finally, the study provides an illustration of a semi-automated brand classification within a specific economic sector and discovers a geographical variation in naming strategies that could be further investigated.

# Data availability

A dataset about Russian law firms, vectors of law firm names obtained from Yandex Foundation models, results of manual labelling, additional files used for mapping, as well as all the code used to conduct the study and generate the paper file are available at Zenodo: https://doi.org/10.5281/zenodo.13691418.

# References

<div id="refs"></div>

```{r fig-1-naming-patterns-and-strategies, fig.cap="**Fig 1.** Dendrogram of law firm naming patterns and strategies in Russia", fig.height=8}
names_classification_plot
```

```{r table-1-clusters}
kable(
  clusters_info_table,
  col.names = c(
    "# of cluster", "# of names", "Description",
    "Pattern", "Strategy"
  ),
  caption = "Table 1. Characteristics of clusters of law firm names"
)
```

```{r fig-2-global-moran-i, fig.cap="Fig. 2. Global Moran's *I* statistics and p-values for law firm naming patterns", fig.height=4}
moran_test_global_res_plot
```

```{r fig-3-localmoran-maps, fig.cap="Fig. 3. Maps of hotspots in the spatial distribution of three law firms naming patterns", fig.height=5}
localmoran_maps
```